---
title: "Simple Linear Regression"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 -  Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
execute:
  freeze: auto
  echo: true
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

## Application exercise

::: appex
📋 [AE 03 - DC Bikeshare](https://deepnote.com/workspace/CofI-MATH-269748c4-4966-4cc3-b064-2a31b9db85a4/project/AE-03-Bikeshare-e02ff833-6768-42e4-8b5d-a62f201324b2/Dockerfile?utm_source=share-modal&utm_medium=product-shared-content&utm_campaign=notebook&utm_content=e02ff833-6768-42e4-8b5d-a62f201324b2)
:::

Complete Exercises 0 and 1.

# Introduction to Simple Linear Regression

## Topics

-   Use simple linear regression to describe the relationship between a quantitative predictor and quantitative response variable.

-   Estimate the slope and intercept of the regression line using the least squares method.

## Computation set up {.smaller}

```{r}
#| echo: true
#| message: false

# load packages
library(tidyverse)       # for data wrangling
library(ggformula)       # for plotting
library(broom)           # for formatting model output
library(knitr)           # for formatting tables

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_bw(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%"
)
```


# Data

## DC Bikeshare

Our data set contains daily rentals from the Capital Bikeshare in Washington, DC in 2011 and 2012. It was obtained from the `dcbikeshare` data set in the `dsbox` R package.

We will focus on the following variables in the analysis:

-   `count`: total bike rentals
-   `temp_orig`: Temperature in degrees Celsius
-   `season`: 1 - winter, 2 - spring, 3 - summer, 4 - fall

Click [here](https://tidyverse.github.io/dsbox/reference/dcbikeshare.html) for the full list of variables and definitions.

Let's complete Exercises 2-6 together

## Data prep {.smaller}

-   Exercise 2: Recode `season` as a factor with names instead of numbers (livecode)
-   Remember:
    +   Think of `|>` as "and then"
    +   `mutate` creates new columns and changes (mutates) existing columns
    +   R calls categorical data "factors"
    
```{r}
bikeshare <- read_csv("../data/dcbikeshare.csv") |> 
  mutate(season = case_when(
    season == 1 ~ "winter",
    season == 2 ~ "spring",
    season == 3 ~ "summer",
    season == 4 ~ "fall"
  ),
  season = factor(season))
```

## Exploratory data analysis (Exercise 3)

```{r}
#| echo: TRUE

gf_point(count ~ temp_orig | season, data = bikeshare) |> 
  gf_labs(x = "Temperature (Celsius)",
          y = "Daily bike rentals")
```

## More data prep

-   (Exercise 5) Filter your data for the season with the strongest relationship and give the resulting data set a new name

```{r}
winter <- bikeshare |> 
  filter(season == "winter")
```

## Rentals vs Temperature

**Goal**: Fit a line to describe the relationship between the temperature and the number of rentals in winter.

```{r}
#| echo: false
#| fig-align: center
p <- ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(x = "Temperature (Celsius)",
       y = "Daily bike rentals")

p
```

## Why fit a line?

We fit a line to accomplish one or both of the following:

. . .

::: {style="color: #799100"}
**Prediction**
:::

How many rentals are expected when it's 10 degrees out?

. . .

::: {style="color : #799100"}
**Inference**
:::

Is temperature a useful predictor of the number of rentals? By how much is the number of rentals expected to change for each degree Celsius?

## Population vs. Sample

**Population:** The set of items or events that you're interested in and hoping (able) to generalize the results of your analysis to.

**Sample:** The set of items that you have data for.

**Representative Sample:** A sample that looks like a small version of your population.

**Goal:** Build a model from your sample which generalizes to your population.


## Terminology

::: columns
::: {.column width="30%"}
-   **Response, *Y***: variable describing the outcome of interest

-   **Predictor, *X***: variable we use to help understand the variability in the response
:::

::: {.column width="70%"}
```{r}
#| echo: false
#| fig-align: center
p
```
:::
:::

## Regression model

**Regression model:** a function that describes the relationship between a quantitative response, $Y$, and the predictor, $X$ (or many predictors).

```{=tex}
\begin{aligned} Y &= \color{black}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{black}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{black}{\boldsymbol{\mu_{Y|X}}} + \epsilon \end{aligned}
```

## Regression model

::: columns
::: {.column width="30%"}
```{=tex}
\begin{aligned} Y &= \color{purple}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{purple}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{purple}{\boldsymbol{\mu_{Y|X}}} + \epsilon \end{aligned}
```
:::

::: {.column width="70%"}
```{r}
#| echo: false
#| fig-align: center
ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "purple", se = FALSE) +
  labs(x = "X" ,
       y = "Y") +
  theme(axis.text = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_blank())
```
:::
:::

$\mu_{Y|X}$ is the mean value of $Y$ given a particular value of $X$.

## Regression model

::: columns
::: {.column width="30%"}
$$
\begin{aligned} Y &= \color{purple}{\textbf{Model}} + \color{blue}{\textbf{Error}} \\[5pt]
&= \color{purple}{\mathbf{f(X)}} + \color{blue}{\boldsymbol{\epsilon}} \\[5pt]
&= \color{purple}{\boldsymbol{\mu_{Y|X}}} + \color{blue}{\boldsymbol{\epsilon}} \\[5pt]
 \end{aligned}
$$
:::

::: {.column width="70%"}
```{r}
#| echo: false
#| fig-align: center
m <- lm(count ~ temp_orig, data = winter)
ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "purple", se = FALSE) +
  geom_segment(aes(x=temp_orig, xend=temp_orig, y=count, yend=predict(m, winter)), color = "blue", linetype = "dotted") +
  labs(x = "X" ,
       y = "Y") +
  theme_bw() +
  theme(axis.text = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_blank())
```
:::
:::

# Simple linear regression (SLR)

## SLR: Statistical model

- **Simple linear regression:** model to describe the relationship between $Y$ and $X$ where:
  + $Y$ is a quantitative/numerical response
  + $X$ is a *single* quantitative predictor
  + $$\Large{Y = \mathbf{\beta_0 + \beta_1 X} + \epsilon}$$

. . .

-   $\beta_1$: True slope of the relationship between $X$ and $Y$
-   $\beta_0$: True intercept of the relationship between $X$ and $Y$
-   $\epsilon$: Error

## SLR: Regression equation

$$\Large{\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X}$$

-   $\hat{\beta}_1$: Estimated slope of the relationship between $X$ and $Y$
-   $\hat{\beta}_0$: Estimated intercept of the relationship between $X$ and $Y$
-   $\hat{Y}$: Predicted value of $Y$ for a given $X$
-   No error term!

## Choosing values for $\hat{\beta}_1$ and $\hat{\beta}_0$

```{r}
#| echo: false
#| fig-align: center
ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.4) +
  geom_abline(intercept = -111.0, slope = 222.4, color = "red") +
  geom_abline(intercept = 1000, slope = 100, color = "green") +
  geom_abline(intercept = -1000, slope = 300, color = "blue") +
  labs(x = "Temperature (Celsius)",
       y = "Daily bike rentals")
```

## Residuals

```{r}
#| echo: false
#| fig-align: center
ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method ="lm", color = "red", se = FALSE) +
  geom_segment(aes(x=temp_orig, xend=temp_orig, y=count, yend=predict(m, winter)), color = "blue", linetype = "dotted") +
  labs(x = "Temperature (Celsius)",
       y = "Daily bike rentals") +
  theme(legend.position = "none")
```

$$\text{residual} = \text{observed} - \text{predicted} = y_i - \hat{y}_i$$

## Least squares line

-  **Residual** for the $i^{th}$ observation:

$$e_i = \text{observed} - \text{predicted}
= y_i - \hat{y}_i$$

-   **Sum of squared residuals**:

$$e^2_1 + e^2_2 + \dots + e^2_n$$

-   **Least squares line** is the one that minimizes the sum of squared residuals

```{r}
#| echo: false
sx <- round(sqrt(var(winter$temp_orig)),4)
sy <- round(sqrt(var(winter$count)),4)
r <- round(cor(winter$temp_orig, winter$count),4)
xbar <- round(mean(winter$temp_orig),4)
ybar <- round(mean(winter$count),4)
```


## Recap

::: {.incremental style="font-size: 0.85em"}

- Simple linear regression (SLR) describes the relationship between a quantitative predictor ($X$) and response ($Y$).
- The regression equation: $\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X$
- $\hat{\beta}_1$ (slope): estimated change in $Y$ for each unit increase in $X$
- $\hat{\beta}_0$ (intercept): estimated value of $Y$ when $X = 0$
- Residuals: difference between observed and predicted values ($e_i = y_i - \hat{y}_i$)
- Least squares method chooses coefficients to minimize the sum of squared residuals
- SLR can be used for prediction and inference
:::

