---
title: "Simple Linear Regression"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 -  Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
execute:
  freeze: auto
  echo: true
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---



# Introduction to Simple Linear Regression

## Topics

-   Interpret the slope and intercept of the regression line.

-   Use R to fit and summarize regression models.

## Computation set up {.smaller}

```{r}
#| echo: true
#| message: false

# load packages
library(tidyverse)       # for data wrangling
library(ggformula)       # for plotting
library(broom)           # for formatting model output
library(knitr)           # for formatting tables

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_bw(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%"
)
```


# Recap

## DC Bikeshare

Our data set contains daily rentals from the Capital Bikeshare in Washington, DC in 2011 and 2012. It was obtained from the `dcbikeshare` data set in the `dsbox` R package.

We will focus on the following variables in the analysis:

-   `count`: total bike rentals
-   `temp_orig`: Temperature in degrees Celsius
-   `season`: 1 - winter, 2 - spring, 3 - summer, 4 - fall

Click [here](https://tidyverse.github.io/dsbox/reference/dcbikeshare.html) for the full list of variables and definitions.

## Data prep {.smaller}

- Converted `season` from 1, 2, 3, 5 into "winter", "spring", "summer", "fall" and convered it into a `factor`
- Created a new dataset called `winter` that contains only the `winter` months
    
```{r}
bikeshare <- read_csv("../data/dcbikeshare.csv") |> 
  mutate(season = case_when(
    season == 1 ~ "winter",
    season == 2 ~ "spring",
    season == 3 ~ "summer",
    season == 4 ~ "fall"
  ),
  season = factor(season)) 

winter <- bikeshare |>
  filter(season == "winter")
```

## Rentals vs Temperature

**Goal**: Fit a line to describe the relationship between the temperature and the number of rentals in winter.

```{r}
#| echo: false
#| fig-align: center
p <- ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(x = "Temperature (Celsius)",
       y = "Daily bike rentals")

p
```

## Residuals

```{r}
#| echo: false
#| fig-align: center

m <- lm(count ~ temp_orig, data = winter)
ggplot(data = winter, mapping = aes(x = temp_orig, y = count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method ="lm", color = "red", se = FALSE) +
  geom_segment(aes(x=temp_orig, xend=temp_orig, y=count, yend=predict(m, winter)), color = "blue", linetype = "dotted") +
  labs(x = "Temperature (Celsius)",
       y = "Daily bike rentals") +
  theme(legend.position = "none")
```

$$\text{residual} = \text{observed} - \text{predicted} = y_i - \hat{y}_i$$

## Last Time {.smaller}

::: {.incremental style="font-size: 0.85em"}
- Sample: data that you have
- Population: group you trying to generalize to
- Representative sample: when your sample looks like small version of your population, if not, we say it's a **biased** sample
- Simple linear regression (SLR): assume relationship betweenquantitative predictor ($X$) and response ($Y$) is a line:
  - Statistical model: $Y = \mathbf{\beta_0 + \beta_1 X} + \epsilon$
- The regression equation: $\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X$
  - $\hat{\beta}_1$ (slope): estimated change in $Y$ for each unit increase in $X$
  - $\hat{\beta}_0$ (intercept): estimated value of $Y$ when $X = 0$
- Residuals: difference between observed and predicted values ($e_i = y_i - \hat{y}_i$)
- Least squares method chooses coefficients to minimize the sum of squared residuals
- SLR can be used for prediction and inference
:::

# Slope and intercept

## Properties of least squares regression {.smaller}

::: incremental
-   Passes through center of mass point, the coordinates corresponding to average $X$ and average $Y$: $\hat{\beta}_0 = \bar{Y} - \hat{\beta}_1\bar{X}$

-   Slope has same sign as the correlation coefficient: $\hat{\beta}_1 = r \frac{s_Y}{s_X}$

    + $r$: correlation coefficient
    + $s_Y, s_X$: sample standard deviations of $X$ and $Y$

-   Sum of the residuals is zero: $\sum_{i = 1}^n e_i \approx 0$

    +   Intuition: Residuals are "balanced"

-   The residuals and $X$ values are uncorrelated
:::

## Estimating the slope

$$\large{\hat{\beta}_1 = r \frac{s_Y}{s_X}}$$

::: columns
::: {.column width="50%"}
```{=tex}
\begin{aligned}
s_X &= 4.2121 \\
s_Y &= 1399.942 \\
r &= 0.6692
\end{aligned}
```
:::

::: {.column width="50%"}
```{=tex}
\begin{aligned}
\hat{\beta}_1 &= 0.6692 \times \frac{1399.942}{4.2121} \\
&= 222.417\end{aligned}
```
:::
:::

<br>

::: small
[Click here](https://introregression.netlify.app/98-appendix) for details on deriving the equations for slope and intercept which is easy if you know multivariate calculus.
:::

## Estimating the intercept

$$\large{\hat{\beta}_0 = \bar{Y} - \hat{\beta}_1\bar{X}}$$

::: columns
::: {.column width="40%"}
```{=tex}
\begin{aligned}
&\bar{x} = 12.2076 \\
&\bar{y} = 2604.133 \\
&\hat{\beta}_1 = 222.4167
\end{aligned}
```
:::

::: {.column width="60%"}
```{=tex}
\begin{aligned}\hat{\beta}_0 &= 2604.133 - 222.4167 \times 12.2076 \\
&= -111.0411
\end{aligned}
```
:::
:::

<br>

::: small
[Click here](https://introregression.netlify.app/98-appendix) for details on deriving the equations for slope and intercept.
:::

## Interpretation

-  Slope: For each additional unit of $X$ we expect $Y$ to increase by $\hat{\beta}_1$, on average.
-  Intercept: If $X$ were 0, we predict $Y$ to be $\hat{\beta}_0$


## Does it make sense to interpret the intercept?

`r emo::ji("white_check_mark")` **The intercept is meaningful in the context of the data if**

-   the predictor can feasibly take values equal to or near zero, or

-   there are values near zero in the observed data.

. . .

`r emo::ji("stop_sign")` Otherwise, the intercept may not be meaningful!

-   Note meaningful doesn't mean wrong

## Estimating the regression line in R

-   Let's complete Exercises 7-11

## Fit model & estimate parameters

```{r}
#| echo: true
#| code-line-numbers: "|1"

winter_fit <- lm(count ~ temp_orig, data = winter)
winter_fit
```

## Look at the regression output

```{r}
#| echo: true
#| code-line-numbers: "|2"

winter_fit <- lm(count ~ temp_orig, data = winter)
winter_fit
```


$$\widehat{\text{count}} = -111.0 + 222.4 \times \text{temp_orig}$$

. . .

::: smallest
**Note:** The intercept is off by a tiny bit from the hand-calculated intercept, this is just due to rounding in the hand calculation.
:::

## The regression output

We'll focus on the first column for now...

```{r}
#| echo: true
#| code-line-numbers: "|2"

winter_fit |> 
  tidy() 
```

## Visualize Model

```{r}
#| echo: true

winter |> 
  gf_point(count ~ temp_orig) |> 
  gf_lm()
```

# Prediction

## Our Model

```{=tex}
\begin{aligned}
\widehat{Y} &=  -111.0 + 222.4 \times X\\

\widehat{\text{count}} &=  -111.0 + 222.4 \times \text{temp_orig}
\end{aligned}
```

## Making a prediction

Suppose that it's 15 degrees Celsius outside. According to this model, how many bike rentals should we expect if it's winter?

```{=tex}
\begin{aligned}
\widehat{\text{count}} &= -111.0 + 222.4 \times \text{temp_orig} \\
&= -111.0 + 222.4 \times 15 \\
&= 3225
\end{aligned}
```

## Prediction in R

```{r}
#| echo: true
#| code-line-numbers: "|2|5"

# create a data frame for a new temperature
new_day <- tibble(temp_orig = 15)

# predict the outcome for a new day
predict(winter_fit, new_day)
```

## Exercise 12

In your Deepnote project answer the following question (you'll need to create a new markdown block):

> Using only addition, subtraction, multiplication, and division. Compute the number of rentals our model would predict if it were 10 degrees Celcius outside.


## Exercise 13

> Verify your answer using the `predict` function.


## Recap

::: {.incremental style="font-size: 0.85em"}
-   Used simple linear regression to describe the relationship between a quantitative predictor and quantitative response variable.

-   Used the least squares method to estimate the slope and intercept.

-   Interpreted the slope and intercept.

    -   **Slope**: For every one unit increase in $x$, we expect y to change by $\hat{\beta}_1$ units, on average.
    -   **Intercept**: If $x$ is 0, then we expect $y$ to be $\hat{\beta}_0$ units

-   Predicted the response given a value of the predictor variable.

-   Used `lm` and the `broom` package to fit and summarize regression models in R.
:::

