---
title: "Multiple linear regression (MLR)"
author: "Prof. Eric Friedlander"
footer: "[🔗 MAT 212 - Winter 2025 -  Schedule](https://mat212wi25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

## Trying something new {.smaller}

::: appex
📋 [AE 07 - Multiple Linear Regression and Categorical Predictors](/ae/ae-07-mlr.qmd)
:::

Complete Exercise 0-3.

## Computational setup

```{r packages}
#| echo: true
#| message: false

# load packages
library(tidyverse)
library(broom)
library(mosaic)
library(ISLR2)
library(patchwork)
library(knitr)
library(coursekata)
library(kableExtra)
library(scales)

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# Create new variable

Credit <- Credit |> 
  mutate(Has_Balance = factor(ifelse(Balance == 0, "No", "Yes")))
```

# Considering multiple variables

## Data: Credit Cards

The data is from the `Credit` data set in the **ISLR2** R package. It is a simulated data set of 400 credit card customers.

```{r}
#| echo: false
Credit |>  glimpse()
```

## Variables

**Features (another name for predictors)**

-   `Cards`: Number of credit cards
-   `Rating`: Credit Rating
-   `Has_Balance`: Whether they have a credit card balance

**Outcome**

-   `Limit`: Credit limit

## Outcome: `Limit`

```{r}
#| code-fold: true
Credit |> 
  gf_density(~Limit, fill = "steelblue") |> 
  gf_labs(title = "Distribution of credit limit",
          x = "Credit Limit")|> 
  gf_refine(scale_x_continuous(labels = dollar_format()))
```

```{r}
#| echo: false
favstats(~Limit, data = Credit)|>
  kable("html") |> 
  kable_styling(font_size = 24)
```

## Predictors {.small}

```{r}
#| code-fold: true
#| out.width: "100%"
p1 <- Credit |> 
  gf_density(~Limit, fill = "steelblue") |> 
  gf_labs(title = "Distribution of credit limit",
          x = "Credit Limit")|> 
    gf_refine(scale_x_continuous(labels = dollar_format()))

p2 <- Credit |> 
  gf_histogram(~Rating, binwidth = 50) |> 
  gf_labs(title = "",
       x = "Credit Rating")

p3 <- Credit |> 
  gf_histogram(~Cards, binwidth = 1) |> 
  gf_labs(title = "",
       x = "Number of Credit Cards")

p4 <- Credit |> 
  gf_bar(~Has_Balance)|> 
  gf_labs(title = "",
       x = "Has a Credit Card Balance")

(p1 + p2) / (p3 + p4)
```

## Outcome vs. predictors {.small}

```{r}
#| code-fold: true
library(GGally)

Credit |> 
  select(Limit, Rating, Cards, Has_Balance) |> 
  ggpairs()
```

## Single vs. multiple predictors

So far we've used a *single predictor variable* to understand variation in a quantitative response variable

. . .

Now we want to use *multiple predictor variables* to understand variation in a quantitative response variable

# Multiple linear regression

## Multiple linear regression (MLR) {.smaller}

Based on the analysis goals, we will use a **multiple linear regression** model of the following form

$$
\begin{aligned}\hat{\text{Limit}} ~ = 
\hat{\beta}_0 & + \hat{\beta}_1 \text{Rating} + \hat{\beta}_2 \text{Cards}
\end{aligned}
$$

Similar to simple linear regression, this model assumes that at each combination of the predictor variables, the values of `Limit` follow a Normal distribution.

## Multiple linear regression

**Recall:** The simple linear regression model assumes

$$
Y|X\sim N(\beta_0 + \beta_1 X, \sigma_{\epsilon}^2)
$$

. . .

**Similarly:** The multiple linear regression model assumes

$$
Y|X_1, X_2, \ldots, X_p \sim N(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \dots + \beta_p X_p, \sigma_{\epsilon}^2)
$$

## Multiple linear regression

At any combination of the predictors, the mean value of the response $Y$, is

$$
\mu_{Y|X_1, \ldots, X_p} = \beta_0 + \beta_1 X_{1} + \beta_2 X_2 + \dots + \beta_p X_p
$$

. . .

Using multiple linear regression, we can estimate the mean response for any combination of predictors

$$
\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X_{1} + \hat{\beta}_2 X_2 + \dots + \hat{\beta}_p X_{p}
$$


## Model fit

```{r}
#| echo: false
lim_fit <- lm(Limit ~ Rating   + Cards,
      data = Credit)

tidy(lim_fit) |>
  kable()
```

## Model equation

$$
\begin{align}\hat{\text{Limit}} =  -341.159 &+14.906 \times \text{Rating}\\
& -72.318 \times \text{Cards}
\end{align}
$$

## Interpreting $\hat{\beta}_j$

-   The estimated coefficient $\hat{\beta}_j$ is the expected change in the mean of $y$ when $x_j$ increases by one unit, *holding the values of all other predictor variables constant*.

. . .

::: appex
Complete Exercises 4-6.
:::

## Prediction

::: question
What is the predicted credit limit for a borrower with these characteristics?
:::

```{r}
head(Credit, n = 1) |> kable()
```



<br>

```{r}
#| echo: true

-341.159 +14.906 * 283 -72.318 * 2
```

. . .

The predicted credit limit for an borrower with an credit rating of 700 and who has 2 credit cards is \$`r round(-341.159 +14.906 * 283 -72.318 * 2)`.

## Prediction, revisited {.midi}

Just like with simple linear regression, we can use the `predict` function in R to calculate the appropriate intervals for our predicted values:

```{r}
#| echo: false

new_borrower <- tibble(
  Rating = 283, 
  Cards = 2
)

predict(lim_fit, new_borrower, interval="prediction")
```

::: callout-note
Difference in predicted value due to rounding the coefficients on the previous slide.
:::

Complete Exercise 7.

## Prediction interval for $\hat{y}$

::: question
Calculate a 90% confidence interval for the **predicted credit limit** for an individual borrower an  credit rating of 700, and who has 2 credit cards.
:::

<br>

```{r}
#| echo: true

predict(lim_fit, new_borrower, interval = "prediction", level = 0.90)
```

::: question
When would you use `"confidence"`? Would the interval be wider or narrower?
:::

## Cautions

-   **Do not extrapolate!** Because there are multiple predictor variables, there is the potential to extrapolate in many directions
-   The multiple regression model only shows **association, not causality**
    -   To show causality, you must have a carefully designed experiment or carefully account for confounding variables in an observational study
    
# Categorical Predictors
    
## Indicator variables

-   Suppose there is a categorical variable with $K$ categories (levels)

-   We can make $K$ indicator variables - one indicator for each category

-   An **indicator variable** takes values 1 or 0

    -   1 if the observation belongs to that category
    -   0 if the observation does not belong to that category

## Data manipulation: Create indicator variables for `Has_Balance`

```{r}
#| echo: true

Credit_dummy <- Credit |>
  mutate(
    No_Balance = if_else(Has_Balance == "No", 1, 0),
    Yes_Balance = if_else(Has_Balance == "Yes", 1, 0)
  )

Credit_dummy |>
  select(Has_Balance, No_Balance, Yes_Balance) |>
  slice(1, 12)
```

## Indicators in the model {.smaller}

-   We will use $K-1$ of the indicator variables in the model.
-   The **reference level** or **baseline** is the category that doesn't have a term in the model.
-   The coefficients of the indicator variables in the model are interpreted as the expected change in the response compared to the baseline, holding all other variables constant.
-   This approach is also called **dummy coding** and *R will do this for you*

```{r}
Credit_dummy |>
  select(Has_Balance, No_Balance, Yes_Balance) |>
  slice(1, 12)
```

## Interpreting Categorical Predictors {.smaller}

::::{.columns}
:::{.column}
```{r}
#| echo: false
cat_mod <- lm(Limit ~ Has_Balance, data = Credit)
cat_mod |>
  tidy() |>
  kable(digits = 3)
```
:::
:::{.column}
```{r}
#| echo: false
Credit |>
  gf_jitter(Limit ~ Has_Balance, height = 0, width = 0.15) |>
  gf_model(cat_mod)
```
:::
::::

::: question
-   Where do we see each of the estimates in the plot?
-   Where do we see the values we'd predict in the plot?
-   Are `Has_Balance` and `Limit` **correlated**?
:::


```{r}
#| echo: false

countdown::countdown(minutes = 3, seconds = 0)
```

. . .

Complete Exercises 8-10.

## Model equation

$$
\begin{align}\hat{\text{Limit}} =  2152.722 &+ 3332.746 \times \text{Yes_Balance}
\end{align}
$$

# Adding in another predictor

::: appex
Complete Exercise 11.
:::

## Credit Limit vs. Cards: parallel slopes

```{r}
#| code-fold: true

cards_model <- lm(Limit ~ Cards, data = Credit)

cards_hasbal_model <- lm(Limit ~ Cards + Has_Balance, data = Credit)

p1 <- plotModel(cards_model) |>
  gf_labs(title = "SLR")
p2 <- plotModel(cards_hasbal_model) |>
  gf_labs(title = "Balance Indicator")

p1 + p2
```

## Parallel slopes interpretation {.smaller}

```{r}
#| echo: FALSE

cards_hasbal_model |>
  tidy() |>
  kable(digits = 3)
```

-   Slope of `Cards` is -36.964 regardless of `Has_Balance` level
-   Change in `Has_Balance` corresponds to a shift in the intercept
    +   Intercept for `No` is 2257.044
    +   For `Yes` shift intercept up 3339.198
        -   (i.e. intercept $= 2257.044 + 3339.198 = 5596.242$)
-   Complete Exercise 12
        
# Interaction terms

## Interaction terms {.smaller}

-   Sometimes the relationship between a predictor variable and the response depends on the value of another predictor variable.
-   This is an **interaction effect**.
-   To account for this, we can include **interaction terms** in the model.
-   We want a model of the form:

$$
\begin{aligned}\hat{Y} ~ = 
\hat{\beta}_0 & + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 +
\hat{\beta}_3X_1\times X_2
\end{aligned}
$$

## Interaction terms: Quantitative and Categorical

$$
\begin{aligned}\hat{\text{Limit}} ~ = 
\hat{\beta}_0 & + \hat{\beta}_1 \text{Cards} + \hat{\beta}_2 \text{Has_Balance} +
\hat{\beta}_3\text{Cards}\times\text{Has_Balance}
\end{aligned}
$$


## Interest rate vs. cards: interaction term

```{r}
#| echo: false

cards_hasbal_interaction_model <- lm(Limit ~ Cards + Has_Balance + Has_Balance*Cards, data = Credit)

p1 <- plotModel(cards_hasbal_model) |>
  gf_labs(title = "No Interaction Term")
p2 <- plotModel(cards_hasbal_interaction_model) |>
  gf_labs(title = "Interaction Term")

p1 + p2+ plot_layout(guides = "collect")
```

## Bike Rentals vs. Temperature: interaction term

```{r}
#| echo: false

bikeshare <- read_csv("../data/dcbikeshare.csv") |> 
  mutate(season = case_when(
    season == 1 ~ "winter",
    season == 2 ~ "spring",
    season == 3 ~ "summer",
    season == 4 ~ "fall"
  ),
  season = factor(season))

bike_model <- lm(count ~ temp_orig + season, data = bikeshare)
bike_model_interaction <- lm(count ~ temp_orig + season + temp_orig*season, data = bikeshare)

p1 <- plotModel(bike_model) |>
  gf_labs(title = "No Interaction Term")
p2 <- plotModel(bike_model_interaction) |>
  gf_labs(title = "Interaction Term")

p1 + p2+ plot_layout(guides = "collect")
```


## Interpreting interaction terms {.smaller}

```{r}
#| echo: FALSE

cards_hasbal_interaction_model |>
  tidy() |>
  kable(digits = 3)
```

-   Slope of `Rating` **depends** on `Has_Balance` level
-   **Equivalent:** fit two separate linear models on the data corresponding to each level of `Has_Balance`

## Understanding the model {.midi}

$$
\begin{aligned}
\hat{Limit} &= 2324.664 - 60.924 \times Cards \\ 
&\qquad+ 3257.976 \times Has\_Balance\\
&\qquad+ 28.499 \times Cards \times Has\_Balance
\end{aligned}
$$

## Interpreting the interaction term

-  For a borrower in no balance, the slope of `Cards` is $-60.924$
-  For a borrower with a balance, the slope of `Cards` is $(-60.924 + 28.499=-32.425)$

::: appex
Complete Exercise 13 and 14.
:::

## Interaction terms: Two Quantitative

$$
\begin{aligned}\hat{\text{Limit}} ~ = 
\hat{\beta}_0 & + \hat{\beta}_1 \text{Cards} + \hat{\beta}_2 \text{Rating} +
\hat{\beta}_3\text{Cards}\times\text{Rating}
\end{aligned}
$$

## Interpreting interaction terms

-   What the interaction means: The effect of the number of open credit cards on the credit limit depends on the borrowers credit rating and vice versa

## Visualizing Model: No Interaction

```{r}
#| echo: FALSE
library(plotly)
library(reshape2)
credit_plot <- plot_ly(Credit, 
         x = ~Rating, 
         y = ~Cards, 
         z = ~Limit,
         color = I("red"),
         opacity = 1,
         type = "scatter3d", 
         mode = "markers",
         size = 1)

axis_x <- seq(min(Credit$Rating), max(Credit$Rating), length.out = 100)
axis_y <- seq(min(Credit$Cards), max(Credit$Cards), length.out = 100)

#Sample points
credit_lm_surface <- expand.grid(Rating = axis_x,Cards = axis_y,KEEP.OUT.ATTRS = F)
credit_lm_surface$Limit <- predict.lm(lim_fit, newdata = credit_lm_surface)
credit_lm_surface <- acast(credit_lm_surface, Cards ~ Rating, value.var = "Limit")

credit_plot |> 
  add_surface(z = credit_lm_surface,
           x = axis_x,
           y = axis_y,
           opacity = 0.8,
           inherit = FALSE,
           contours = list(z = list(show = TRUE, width = 2,
                                       highlightcolor = "#ff0000", 
                                       project = list(z = TRUE)),
                              coloring = "heatmap"))
```


## Visualizing Interaction Model: Exaggerated

```{r}
#| echo: FALSE

Credit_sim <- tibble(Cards = round(runif(100, 0, 20)), Rating = runif(100, 0, 1000), 
                     Limit = 100 + 50*Cards + Rating + 1000*Cards*Rating)

credit_plot_sim <- plot_ly(Credit_sim, 
         x = ~Rating, 
         y = ~Cards, 
         z = ~Limit,
         color = I("red"),
         opacity = 1,
         type = "scatter3d", 
         mode = "markers",
         size = 1)

axis_x_sim <- seq(min(Credit_sim$Rating), max(Credit_sim$Rating), length.out = 100)
axis_y_sim <- seq(min(Credit_sim$Cards), max(Credit_sim$Cards), length.out = 100)

lim_fit_int_sim <- lm(Limit ~ Rating  + Cards + Rating * Cards,
      data = Credit_sim)

#Sample points
credit_lm_surface <- expand.grid(Rating = axis_x_sim, Cards = axis_y_sim,KEEP.OUT.ATTRS = F)
credit_lm_surface$Limit <- predict.lm(lim_fit_int_sim, newdata = credit_lm_surface)
credit_lm_surface <- acast(credit_lm_surface, Cards ~ Rating, value.var = "Limit")

credit_plot_sim |>
  add_surface(z = credit_lm_surface,
           x = axis_x_sim,
           y = axis_y_sim,
           opacity = 0.8,
           inherit = FALSE,
           contours = list(z = list(show = TRUE, width = 2,
                                       highlightcolor = "#ff0000",
                                       project = list(z = TRUE)),
                              coloring = "heatmap"))
```

## Visualizing Interaction Model: Real

```{r}
#| echo: FALSE

lim_fit_int <- lm(Limit ~ Rating  + Cards + Rating * Cards,
      data = Credit)

#Sample points
credit_lm_surface <- expand.grid(Rating = axis_x,Cards = axis_y,KEEP.OUT.ATTRS = F)
credit_lm_surface$Limit <- predict.lm(lim_fit_int, newdata = credit_lm_surface)
credit_lm_surface <- acast(credit_lm_surface, Cards ~ Rating, value.var = "Limit")

credit_plot |>
  add_surface(z = credit_lm_surface,
           x = axis_x,
           y = axis_y,
           opacity = 0.8,
           inherit = FALSE,
           contours = list(z = list(show = TRUE, width = 2,
                                       highlightcolor = "#ff0000",
                                       project = list(z = TRUE)),
                              coloring = "heatmap"))
```


## Model Fit {.smaller}

```{r}
#| echo: FALSE
lm(Limit ~ Cards + Rating + Cards * Rating, data = Credit) |>
  tidy() |>
  kable(digits = 5)
```

$$
\begin{aligned}\hat{\text{Limit}} ~ =
&  -315.005 + 14.834~\text{Rating} -81.116~\text{Cards}\\
& \qquad+ 0.024~\text{Rating}\times\text{Cards}
\end{aligned}
$$

## Interpreting the interaction term

-  For a fixed `Rating` the slope of `Cards` is $(-81.116 + 0.024\times\text{Rating})$
-  For a fixed `Cards` the slope of `Rating` is $(14.834 + 0.024\times\text{Cards})$



## What's actually happening: {.smaller}

```{r}
#| echo: TRUE
Credit_int <- Credit |>
  mutate(Interaction = Cards * Rating)

Credit_int |>
  select(Limit, Cards, Rating, Interaction) |>
  head() |>
  kable()
```

## What's actually happening:

```{r}
lm(Limit ~ Cards + Rating + Interaction, data = Credit_int) |>
  tidy() |>
  kable(digits = 5)
```


# Wrap up

## Recap

-   Introduced multiple linear regression

-   Interpreted coefficients in the multiple linear regression model

-   Calculated predictions and associated intervals for multiple linear regression models

-   Introduced categorical variables

-   Used interaction terms

