---
title: "Model Evaluation"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 - Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---
```{r setup}
#| include: false

library(countdown)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%",
  fig.align = "center"
)
```

# Application exercise

::: appex
📋 AE 10 - Model Evaluation: Open in Deepnote
:::

Complete Exercise 0.

## Computational set up

```{r packages}
#| echo: true
#| message: false

# load packages
library(tidyverse)   # for data wrangling and visualization
library(ggformula)   # for plotting using formulas
library(broom)       # for formatting model output
library(scales)      # for pretty axis labels
library(knitr)       # for pretty tables
library(kableExtra)  # also for pretty tables
library(patchwork)   # arrange plots

# Spotify Dataset
spotify <- read_csv("../data/spotify-popular.csv")

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_bw(base_size = 20))
```

## Quick Data Cleaning

```{r}
#| echo: true

spotify <- spotify |> 
  mutate(duration_min = duration_ms / 60000)
```

::: question
- What is this code doing?
- Why might I be doing it?
:::

## The regression model, revisited {.smaller}

```{r}
#| echo: true
spotify_fit <- lm(danceability ~ duration_min, data = spotify)

tidy(spotify_fit, conf.int = TRUE, conf.level = 0.95) |>
   kable(digits = 3)
```

- There is strong statistical evidence that there is a linear relationship between the duration of a song and it's danceability.

- We are 95% confidence that as the length of a song increases by one minute the danceability will decrease by between 0.009 and 0.039 units.

# Model evaluation

## Partitioning Variability

Let's think about variation:

:::{.incremental}
- DATA = MODEL + ERROR
- $\substack{\text{Variation} \\ \text{in Y}} = \substack{\text{Variation explained} \\ \text{by model}} + \substack{\text{Variation not explained} \\ \text{by model}}$
:::

## Partitioning Variability (ANOVA)

:::{.incremental}
-   $y_i - \bar{y} = (\hat{y}_i - \bar{y}) + (y_i-\hat{y}_i)$
-   Square and sum: $\sum(y_i-\bar{y})^2 = \sum(\hat{y} - \bar{y})^2 + \sum(y-\hat{y})^2$
-   $\substack{\text{Sum of squares} \\ \text{Total}} = \substack{\text{Sum of squares} \\ \text{model}} + \substack{\text{Sum of squares} \\ \text{error}}$
-   $SSTotal = SSModel + SSE$
-   $SST = SSM + SSE$
-   ANOVA: Analysis of Variance
:::

## ANOVA in R

```{r}
spotify_fit |>
  anova() |>
  tidy() |>
  kable() # Ignore this line when working in deepnote
```

:::{.fragment}
- More on this later in the semester
- Complete Exercise 1.
:::

## Recall: Correlation Coefficient

-   The **correlation coefficient**, $r$, is a number between -1 and +1 that measures how strong the linear relationship between two variables $x$ and $y$ is.

$$
r = \frac{\sum(x_i - \bar{x})(y_i-\bar{y})}{\sqrt{\sum(x_i-\bar{x})^2\sum(y_i-\bar{y})^2}}
= \frac{\sum(x_i - \bar{x})(y_i-\bar{y})}{s_xs_y}
$$


## Two statistics: $R^2$ {.smaller}

::: incremental
-   **R-squared**, $R^2$, **Coefficient of Determination** : Percentage of variability in the outcome explained by the regression model (in the context of SLR, the predictor)
$$
R^2 = Cor(y, \hat{y})^2
$$
    -  Also called **PRE (Percent Reduction in Error)** because:
$$
R^2 = \frac{SSModel}{SSTotal}
$$
:::


## Two statistics: RMSE {.smaller}

-   **Root mean square error, RMSE**: A measure of the average error (average difference between observed and predicted values of the outcome)
$$
RMSE = \sqrt{\frac{\sum_{i = 1}^n (y_i - \hat{y}_i)^2}{n}}
$$
    + Sometimes people just care about numerator (SSE) or version without the square-root (MSE)
    + Sometimes the denominator may have $n-1$ instead

. . .

::: question
What indicates a good model fit? Higher or lower $R^2$? Higher or lower RMSE?
:::

## $R^2$

::: incremental
-   Ranges between 0 (terrible predictor) and 1 (perfect predictor)
-   Has no units
-   Calculate with `rsq()` from `yardstick` package using the augmented data:
```{r}
#| echo: true

library(yardstick)
spotify_aug <- augment(spotify_fit)

rsq(spotify_aug, truth = danceability, estimate = .fitted) |> kable()

```
:::

## Interpreting $R^2$ {.smaller}

```{r}
#| echo: false
#|
spotify_fit_rsq <- round(rsq(spotify_aug, truth = danceability, estimate = .fitted) |> pull(.estimate) * 100, 1)
```

::: poll
🗳️ **Discussion**

::: midi
::: poll
The $R^2$ of the model for `danceability` from `Average_Income_K` is `r spotify_fit_rsq`%. Which of the following is the correct interpretation of this value?
:::

1.  `duration_min` correctly predicts `r spotify_fit_rsq`% of `danceability`.
2.  `r spotify_fit_rsq`% of the variability in `danceability` can be explained by `duration_min`.
3.  `r spotify_fit_rsq`% of the variability in `duration_min` can be explained by `danceability`.
4.  `r spotify_fit_rsq`% of the time `danceability` can be predicted by `duration_min`.
:::
:::

Complete Exercise 2.

## Activity

::: appex
In groups, at the board, design a simulation-based procedure for producing a p-value for the following hypothesis test.

- $H_0: R^2 = 0$
- $H_A: R^2 \neq 0$
:::

. . .

Complete Exercise 3.

## RMSE {.smaller}

::: incremental
-   Ranges between 0 (perfect predictor) and infinity (terrible predictor)

-   Same units as the response variable

-   Interpretation (kind of): how much does my model miss by, on average.

-   Calculate with  `rmse()` from `yardstick` package using the augmented data:

```{r}
#| echo: true

rmse(spotify_aug, truth = danceability, estimate = .fitted)
```
:::

-   Complete Exercise 4.

## Using the word "Good"

-   There is no such thing as a "Good" $R^2$ or, especially, RMSE without context
-   Whether your model is a "Good" model depends on many things:
    +   What are you using your model for?
    +   How good are other models?

## Recap {.smaller}

-   Can decompose total variation (SST) into variation explained by the model (SSM) and leftover variation (SSE)
-   Two metrics for evaluating and comparing models: 
    +   $R^2$: What proportion of the variation in the response variable is explained by the model?
    +   $RMSE$: How far is does my model miss by on average?


