---
title: "SLR: Transformations"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 - Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

```{r setup}
#| include: false

library(countdown)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%",
  fig.align = "center"
)
```

## Application exercise

::: appex
📋 AE 10/11 - Transformations + Outliers: Complete in Deepnote
:::

Complete Exercises 0-3.

## Computational set up

```{r packages}
#| echo: true
#| message: false

# load packages
library(tidyverse)   # for data wrangling and visualization
library(broom)       # for formatting model output
library(ggformula)   # for creating plots using formulas
library(scales)      # for pretty axis labels
library(knitr)       # for pretty tables
library(moderndive)  # for house_price dataset
library(fivethirtyeight)   # for fandango dataset
library(kableExtra)  # also for pretty tables
library(patchwork)   # arrange plots

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_bw(base_size = 20))
```

## Data: `house_prices` {.smaller}

::: nonincremental
-   Contains house sale prices for King County, which includes Seattle, from homes
sold between May 2014 and May 2015
-   Obtained from [Kaggle.com](https://www.kaggle.com/harlfoxem/housesalesprediction/data)
-   Imported from the `moderndive` package
:::

```{r}
glimpse(house_prices)
```

## Variables

- Outcome
  + `price`: the sale price in dollars
- Predictor 
  + `sqft_living`: the square footage of the home
  
## Recap: Fit the model {.smaller}

```{r}
hp_fit <- lm(price ~ sqft_living, data = house_prices)
tidy(hp_fit) |>  kable()
```

:::{.incremental}
- Write down the model:
  + **Model:** $\hat{\text{price}} = -43580.74 + 280.62\times\text{sqft_living}$
- Interpret the slope and intercept in the context of this problem:
  + **Interpretation:** If the square footage of the house increases by 1, the price increases by and average of \$280.62 and a (theoretical) house with 0 square feet with cost -\$43,580.74.
:::

## Recap: Fit the model

```{r}
#| code-fold: true

gf_point(price ~ sqft_living, data = house_prices,
         alpha = 0.25, size = 0.01) |> 
  gf_smooth(method = "lm", color = "red") |> 
  gf_labs(x = "Square Footage", 
       y = "Sale Price")  |> 
  gf_refine(scale_y_continuous(labels = label_dollar()),
  scale_x_continuous(labels = label_number()))
```

## Recap: Model conditions

1.  **Linearity:** There is a linear relationship between the outcome and predictor variables
2.  **Constant variance:** The variability of the errors is equal for all values of the predictor variable
3.  **Normality:** The errors follow a normal distribution
4.  **Independence:** The errors are independent from each other

. . .

::: question
How should we check these assumptions?
:::

## Recap: Residual Histogram

```{r}
#| code-fold: true

hp_aug <- augment(hp_fit)

gf_histogram(~.resid, data = hp_aug, bins = 100) |> 
  gf_labs(x = "Residual", 
       y = "Count", 
       title = "Residual Histogram")
```

## Recap: QQ-Plot of Residuals

```{r}
#| code-fold: true

gf_qq(~.resid, data = hp_aug) |> 
  gf_qqline() |>
  gf_labs(x = "Theoretical quantile", 
       y = "Observed quantile", 
       title = "Normal QQ-plot of residuals")
```

## Recap: Residuals vs. Fitted Values

```{r}
#| code-fold: true
gf_point(.resid ~ .fitted, data = hp_aug, 
         alpha = 0.25, size = 0.01) |> 
  gf_hline(yintercept = 0, linetype = "dashed") |> 
  gf_labs(
    x = "Fitted value", y = "Residual",
    title = "Residuals vs. fitted values"
  )
```


## [Are model conditions satisfied?]{.r-fit-text}

1.  **Linearity:** ```r emo:::ji("question")```
2.  **Constant variance:** ```r emo:::ji("x")```
3.  **Normality:** ```r emo:::ji("x")```
4.  **Independence:** ```r emo:::ji("question")```

## [What to do when regression conditions are violated]{.r-fit-text}

Examples:

1. Lack of normality in residuals 
2. Patterns in residuals
3. Heteroscedasticity (non-constant variance)
4. Outliers: influential points, large residuals


# Transformations

## Data Transformations

Can be used to:

a. Address nonlinear patterns
b. Stabilize variance
c. Remove skewness from residuals
d. Minimize effects of outliers

## Common Transformations

For either the response $Y$ or predictor $X$:

- Logarithm $Z \to \log(Z)$
  - Note: "log" means "log base $e$"
- Square Root $Z \to \sqrt{Z}$
- Exponential $Z \to e^Z$
- Power functions $Z \to Z^2, Z^3, Z^4, \ldots$
- Reciprocal $Z \to 1/Z$

## General Approach

- Fix non-constant variance by transforming $Y$ (do this first)
    + Fan Pattern: Log-Transform $Y$
- Fix non-linearity by transforming $X$


## Why a Log Transformation? {.smaller}

> Some relationship are *multiplicative* (not linear)

Example: Area of a circle

$$
\begin{aligned}
A &= \pi r^2 \text{ (not linear)}\\
\log(A) &= \log(\pi r^2)
= \log(\pi) + 2\log(r)\\
\log(A) &= \beta_0 + \beta_1\times \log(r)\\
\implies & \log(A) \text{ is a linear function of } \log(r)
\end{aligned}
$$

Look for:

- Increasing variability in scatterplot
- Strongly right-skewed residual distributions
- Complete Exercise 4

## Fixing non-linearity

-   Many departures from linearity can be solved with power transformations (e.g. $X^{power}$)

    +   For technical reasons, $power = 0$ corresponds to $\log$

-   Concave down pattern $\Rightarrow$ transform down (i.e. $power < 1$)

    +   $\log$ is typically a good first choice

-   Concave up pattern $\Rightarrow$ transform up (i.e. $power > 1$)

-   Complete Exercises 5-8.


## Back to `house_sales`

```{r}
#| code-fold: true

p1 <- gf_point(price ~ sqft_living, data = house_prices,
         alpha = 0.25, size = 0.01) |> 
  gf_smooth(method = "lm", color = "red") |> 
  gf_labs(x = "Square Footage", 
       y = "Sale Price")  |> 
  gf_refine(scale_y_continuous(labels = label_dollar()),
  scale_x_continuous(labels = label_number()))

p2 <- gf_point(log(price) ~ sqft_living, data = house_prices,
         alpha = 0.25, size = 0.01) |> 
  gf_smooth(method = "lm", color = "red") |> 
  gf_labs(x = "Square Footage", 
       y = "log(Sale Price)")  |> 
  gf_refine(scale_y_continuous(labels = label_dollar()),
  scale_x_continuous(labels = label_number()))

p3 <- gf_point(price ~ log(sqft_living), data = house_prices,
         alpha = 0.25, size = 0.01) |> 
  gf_smooth(method = "lm", color = "red") |> 
  gf_labs(x = "log(Square Footage)", 
       y = "Sale Price")  |> 
  gf_refine(scale_y_continuous(labels = label_dollar()),
  scale_x_continuous(labels = label_number()))

p4 <- gf_point(log(price) ~ log(sqft_living), data = house_prices,
         alpha = 0.25, size = 0.01) |> 
  gf_smooth(method = "lm", color = "red") |> 
  gf_labs(x = "log(Square Footage)", 
       y = "log(Sale Price)")  |> 
  gf_refine(scale_y_continuous(labels = label_dollar()),
  scale_x_continuous(labels = label_number()))

(p1 + p2)/ (p3 + p4)
```

## Fitting Transformed Models {.smaller}

::::{.columns}
:::{.column}
```{r}
logprice_model <- lm(log(price) ~ sqft_living, data = house_prices)
tidy(logprice_model) |> kable()
```

$$
\begin{aligned}
\log(Y) &= 12.22  + 3.99\times 10^{-4}\times X\\
Y &= e^{12.22 + 3.99\times 10^{-4}\times X}\\
&= 202805\times e^{3.99\times 10^{-4}\times X}
\end{aligned}
$$
:::
:::{.column}
```{r}
loglog_model <- lm(log(price) ~ log(sqft_living), data = house_prices)
tidy(loglog_model) |> kable()
```
$$
\begin{aligned}
\log(Y) &=6.73 + 0.837\times \log(X)\\
\log(Y) &= \log(e^{6.73})  + \log(X^{0.837})\\
Y &= 873.15\times X^{0.837}
\end{aligned}
$$
:::
::::

## Residuals Histograms

```{r}
#| code-fold: true

lp_aug <- augment(logprice_model)
ll_aug <- augment(loglog_model)

p1 <- gf_histogram(~.resid, data = lp_aug, bins = 100) |> 
  gf_labs(x = "Residual", 
       y = "Count", 
       title = "Log Price Residuals")

p2 <- gf_histogram(~.resid, data = ll_aug, bins = 100) |> 
  gf_labs(x = "Residual", 
       y = "Count", 
       title = "Log-Log Residuals")

(p1 + p2)
```

## QQ-Plots of Residuals

```{r}
#| code-fold: true

p1 <- gf_qq(~.resid, data = lp_aug) |> 
  gf_qqline() |>
  gf_labs(x = "Theoretical quantile", 
       y = "Observed quantile", 
       title = "Log Price QQ")

p2 <- gf_qq(~.resid, data = ll_aug) |> 
  gf_qqline() |>
  gf_labs(x = "Theoretical quantile", 
       y = "Observed quantile", 
       title = "Log-Log QQ")

p1 + p2
```

## Residuals vs. Fitted Values

```{r}
#| code-fold: true
p1 <- gf_point(.resid ~ .fitted, data = lp_aug, 
         alpha = 0.25, size = 0.01) |> 
  gf_hline(yintercept = 0, linetype = "dashed") |> 
  gf_labs(
    x = "Fitted value", y = "Residual",
    title = "Log Price Model"
  )

p2 <- gf_point(.resid ~ .fitted, data = ll_aug, 
         alpha = 0.25, size = 0.01) |> 
  gf_hline(yintercept = 0, linetype = "dashed") |> 
  gf_labs(
    x = "Fitted value", y = "Residual",
    title = "Log-Log Model"
  )

p1 + p2
```

## A note on evaluation

If you are computing your evaluation metrics (e.g. $R^2$ or RMSE), you should transform your predictions BACK to their original scale, especially if you're trying to choose the best model

:::{.question}
- Why do we need to undo the transformation for evaluation metrics by not residuals plots? 
- Why don't we need to worry about the predictors?
:::

Complete Exercise 9.

## Recap {.smaller}

- Transformations
  + Transform $Y$ to fix non-constant variance (and non-normality)
  + Transform $X$ to fix non-linearity
  + Power transformations are powerful (concave up/drown $\Rightarrow$ power up/down)
  + logs allow us to model a lot of non-linear relationships with a linear model