<<<<<<< HEAD
---
title: "MLR: Inference and Conditions"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 - Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

```{r}
#| include: false

# figure options
knitr::opts_chunk$set(
  fig.width = 10, fig.asp = 0.618,
  fig.retina = 3, dpi = 300, fig.align = "center"
)
```

## Application Exercise

::: appex
📋 AE 15 - Inference for MLR

- Complete Exercises 0-3.
:::


## Topics

::: nonincremental
-   Inference for multiple linear regression
:::

## Computational setup

```{r}
#| echo: true


# load packages
library(tidyverse)
library(broom)
library(mosaic)
library(mosaicData)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)
library(GGally)
library(countdown)
library(rms)
library(Stat2Data)
library(coursekata)

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))
```


## Data: `Tadpools` {.smaller}

::: nonincremental
Biologists wondered whether tadpoles can adjust the relative length of their intestines if they are exposed to a fungus called Batrachochytrium dendrobatidis (Bd).
:::

```{r}
#| echo: false
data("Tadpoles")
glimpse(Tadpoles)
```


Source: [Stat2Data Package](https://cran.r-project.org/web/packages/Stat2Data/refman/Stat2Data.html#Tadpoles).

## Visualizing the data

```{r}
Tadpoles |> ggpairs()
```


# Conduct a hypothesis test for $\beta_j$

## [Review: Simple linear regression (SLR)]{.r-fit-text}

```{r}
#| echo: false
p1 <- gf_jitter(GutLength ~ Treatment , data = Tadpoles, alpha = 0.5, height = 0, width = 0.1) |>
  gf_model(lm(GutLength ~ Treatment, data = Tadpoles))  |>
  gf_labs(x = "Treatment Group", y = "Gut Length (mm)")

p2 <- gf_point(GutLength ~ MouthpartDamage , data = Tadpoles, alpha = 0.5) |>
  gf_model(lm(GutLength ~ MouthpartDamage, data = Tadpoles))  |>
  gf_labs(x = "Mouthpart Damage", y = "Gut Length (mm)")

p1 + p2
```



## SLR model summary

```{r}
#| echo: false
treatment_model <- lm(GutLength ~ Treatment, data = Tadpoles)

tidy(treatment_model) |> kable()

damage_model <- lm(GutLength ~ MouthpartDamage, data = Tadpoles)

tidy(damage_model) |> kable()
```


## SLR hypothesis test {.smaller}

```{r}
#| echo: false

tidy(treatment_model) |> kable()
```

1.  **Set hypotheses:** $H_0: \beta_1 = 0$ vs. $H_A: \beta_1 \ne 0$

. . .

2.  **Calculate test statistic and p-value:** The test statistic is $t= 3.12$ . The p-value is calculated using a $t$ distribution with 88 degrees of freedom. The p-value is $0.005$.

. . .

3.  **State the conclusion:** The p-value is small, so we reject $H_0$. The data provide strong evidence that there exposure to fungus is associated with a decrease in gut length... hmmm...

## Controlling for counfounders

- Controlling for confounders:
  + Body length
  + Mouthpart Damage

```{r}
#| code-fold: hide
p1 <- gf_point(GutLength ~ Body, color = ~Treatment, data = Tadpoles, alpha = 0.5) |>
  gf_labs(x = "Body Length (mm)", y = "Gut Length (mm)")

p2 <- gf_point(GutLength ~ MouthpartDamage , color = ~Treatment, data = Tadpoles, alpha = 0.5)|>
  gf_labs(x = "Mouthpart Damage", y = "Gut Length (mm)")

p1 + p2 + plot_layout(guides = "collect")
```

Discussion: How should we control for our confounders?

## Multiple linear regression

```{r}
Tadpoles <- Tadpoles |>
  mutate(GutProp = GutLength / Body)

treat_damage_model <- lm(GutProp ~ Treatment + MouthpartDamage, data = Tadpoles)

tidy(treat_damage_model) |> kable()
```

## Multiple linear regression

The multiple linear regression model assumes $$Y|X_1, X_2,  \ldots, X_p \sim N(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \dots + \beta_p X_p, \sigma_\epsilon^2)$$

. . .

For a given observation $(x_{i1}, x_{i2}, \ldots, x_{ip}, y_i)$, we can rewrite the previous statement as

$$y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \dots + \beta_p x_{ip} + \epsilon_{i}, \hspace{10mm} \epsilon_i \sim N(0,\sigma_{\epsilon}^2)$$


## Estimating $\sigma_\epsilon$ {.smaller}

For a given observation $(x_{i1}, x_{i2}, \ldots,x_{ip}, y_i)$ the residual is
$$
\begin{aligned}
e_i &= y_{i} - \hat{y_i}\\
&= y_i - (\hat{\beta}_0 + \hat{\beta}_1 x_{i1} + \hat{\beta}_{2} x_{i2} + \dots + \hat{\beta}_p x_{ip})
\end{aligned}
$$

. . .

The estimated value of the regression standard error , $\sigma_{\epsilon}$, is

$$\hat{\sigma}_\epsilon  = \sqrt{\frac{\sum_{i=1}^ne_i^2}{n-p-1}}$$

. . .

As with SLR, we use $\hat{\sigma}_{\epsilon}$ to calculate $SE_{\hat{\beta}_j}$, the standard error of each coefficient. See [Matrix Form of Linear Regression](https://github.com/STA210-Sp19/supplemental-notes/blob/master/regression-basics-matrix.pdf) for more detail.

## MLR hypothesis test: avgtemp {.smaller}

1.  **Set hypotheses:** $H_0: \beta_{avgtemp} = 0$ vs. $H_A: \beta_{avgtemp} \ne 0$, given `hightemp` is in the model

. . .

2.  **Calculate test statistic and p-value:** The test statistic is $t = -3.39$. The p-value is calculated using a $t$ distribution with $$(n - p - 1) = 90 - 2 -1 = 87$$ degrees of freedom. The p-value is $\approx 0.0011$.
  
    + Note that $p$ counts all non-intercept $\beta$'s

. . .

3.  **State the conclusion:** The p-value is small, so we reject $H_0$. The data provides convincing evidence that a avgtemp is a useful predictor *in a model that already contains rail_trail hightemp as a predictor for the rail_trail volume*.


## MLR hypothesis test: interaction terms {.smaller}

- Framework: Same as previous slide except use $\beta$ of interaction term
- $p$ (the number of predictors) should include the interaction term
- Conclusion: tells you whether interaction term is a useful predictor
- Warning: if $X_1$ and $X_2$ have an interaction term, don't try to interpret their individual p-values... if interaction term is significant, then both variables are important

::: appex
Complete Exercises 3 and 4.
:::

# Effect Size and Power

## Effect Size and Power {.smaller}

-   **Effect Size:** 
    +   General Idea: how big is your parameter compared to the variability of your data
    +   Why is it important: if parameter has small effect size, it may not be a useful parameter to include. However, what effect size is useful depends on your problem
    +   In practice: many formal ways of determining effects size

-   **Statistical Power:** the ability of a hypothesis test to detect a given **effect size**
    +   more data = more power
    +   tons of data = tons of power
    +   In practice: think about how what effect size you'd like to be able to detect and design your study to collect enough data to do so

::: appex
Complete Exercise 6
:::


## Effect Size:

-   Cohen's-D (for a categorical variable with two levels like treatment/control): $$d = \frac{|\text{difference in means}|}{\text{standard deviation}} = \frac{|\beta_1|}{\text{standard deviation (not standard error)}}$$
-   R-squared: proportion of variability in response explained by model
-   IMPORT TAKEAWAY:
    -  Effect size is a measure of how important/useful a predictor is in explaining the response variable.
    -  Effect size DOES NOT depend on sample size.
    -  Statistical significance DOES depend on sample size.

## Power

-   Power: probability of rejecting null hypothesis when it is false
-   Factors affecting power:
    +   Effect size: larger effect size = more power
    +   Sample size: larger sample size = more power
    +   Significance level: larger alpha = more power
    +   Variability in data: less variability = more power

# Using Simulation to Think About Power {.smaller}


```{r}
lm(volume ~ season, data = rail_trail) |> tidy()
```

# Confidence interval for $\beta_j$

## Confidence interval for $\beta_j$ {.midi}

-   The $C\%$ confidence interval for $\beta_j$ $$\hat{\beta}_j \pm t^* SE(\hat{\beta}_j)$$ where $t^*$ follows a $t$ distribution with $n - p - 1$ degrees of freedom.

-   **Generically**: We are $C\%$ confident that the interval LB to UB contains the population coefficient of $x_j$.

-   **In context:** We are $C\%$ confident that for every one unit increase in $x_j$, we expect $y$ to change by LB to UB units, *holding all else constant*.

::: appex
Complete Exercise 5.
:::

## Confidence interval for $\beta_j$

```{r}
#| echo: FALSE
tidy(rail_trail_fit, conf.int = TRUE) |>
  kable(digits = 2)
```


# Inference pitfalls

## Large sample sizes

::: callout-caution
If the sample size is large enough, the test will likely result in rejecting $H_0: \beta_j = 0$ even $x_j$ has a very small effect on $y$.

::: nonincremental
-   Consider the **practical significance** of the result not just the statistical significance.

-   Instead of saying **statistically significant** say **statistically detectable**

-   Use confidence intervals to draw conclusions instead of relying only on p-values.
:::
:::

## Small sample sizes

::: callout-caution
If the sample size is small, there may not be enough evidence to reject $H_0: \beta_j=0$.

::: nonincremental
-   When you fail to reject the null hypothesis, **DON'T** immediately conclude that the variable has no association with the response.

-   There may be a linear association that is just not strong enough to detect given your data, or there may be a non-linear association.
:::
:::


# Conditions for inference


## Model conditions

Our model conditions are the same as they were with SLR:

1.  **Linearity:** There is a linear relationship between the response and predictor variables.

2.  **Constant Variance:** The variability about the least squares line is generally constant.

3.  **Normality:** The distribution of the residuals is approximately normal.

4.  **Independence:** The residuals are independent from each other.


## Checking Linearity

-   Look at a plot of the residuals vs. predicted values

-   Look at a plot of the residuals vs. each predictor

    +   Use scatter plots for quantitative and boxplots of categorical predictors

-   Linearity is met if there is no discernible pattern in each of these plots

Complete Exercises 7-10

## Fitting the Full Model {.smaller}

```{r}
full_model <- lm(volume ~ . , data = rail_trail)
full_model |> tidy() |>  kable()
```

## Residuals vs. predicted values

```{r}
#| label: main_res_pred
#| echo: FALSE

rt_full_aug <- augment(full_model)

gf_point(.resid ~ .fitted, data = rt_full_aug, alpha = 0.7) |>
  gf_hline(yintercept = 0, color = "red", linetype = "dashed") |>
  gf_labs(x = "Predicted values", y = "Residuals")
```

::: question
Does the constant variance condition appear to be satisfied?
:::

```{r}
#| ref.label: main_res_pred
#| echo: false
```

## Checking constant variance

-   The vertical spread of the residuals is not constant across the plot.

-   The constant variance condition is not satisfied.

. . .

::: question
Given this conclusion, what might be a next step in the analysis?
:::

## Residuals vs. each predictor

```{r}
#| fig.asp: 0.5
#| echo: false

p1 <- gf_point(.resid ~ hightemp, data = rt_full_aug, alpha = 0.7) |>
  gf_hline(yintercept = 0, color = "red", linetype = "dashed")

p2 <- gf_point(.resid ~ avgtemp, data = rt_full_aug, alpha = 0.7) |>
  gf_hline(yintercept = 0, color = "red", linetype = "dashed")

p3 <- gf_boxplot(.resid ~ season, data = rt_full_aug, alpha = 0.7) |>
  gf_hline(yintercept = 0, color = "red", linetype = "dashed")

p4 <- gf_point(.resid ~ cloudcover, data = rt_full_aug, alpha = 0.7) |>
  gf_hline(yintercept = 0, color = "red", linetype = "dashed")

(p1 + p2) / (p3 + p4)
```

## Checking linearity

-   The plot of the residuals vs. predicted values looked OK

-   The plots of residuals vs. `hightemp` and `avgtemp` appear to have a parabolic pattern.

-   The linearity condition does not appear to be satisfied given these plots.

. . .

::: question
Given this conclusion, what might be a next step in the analysis?
:::

Complete Exercises 11-12.

## Checking normality

```{r}
#| fig.asp: 0.8
#| echo: false

resid_hist <- gf_histogram(~.resid, data = rt_full_aug, binwidth = 25) |> 
  gf_labs(x = "Residuals")

resid_qq <- gf_qq(~.resid, data = rt_full_aug) |>
  gf_qqline()

resid_hist + resid_qq
```

The distribution of the residuals is approximately unimodal and symmetric but the qqplot shows that tails are too heavy, so the normality condition is likely not satisfied. The sample size 90 is sufficiently large to relax this condition.

## Checking independence

-   We can often check the independence condition based on the context of the data and how the observations were collected.

-   If the data were collected in a particular order, examine a scatter plot of the residuals versus order in which the data were collected.

-   If there is a grouping variable lurking in the background, check the residuals based on that grouping variable.

## Recap

- Inference for individual predictors
    + Hypothesis testing: same as SLR but add *if all other terms are in the model*
    + Confidence intervals: same as SLR but add *holding all other variables in the model constant*
- Conditions: Same as for SLR
    + For linearity, also check fitter values vs. each predictor
=======
---
title: "MLR: Inference"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 - Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

```{r}
#| include: false

# figure options
knitr::opts_chunk$set(
  fig.width = 10, fig.asp = 0.618,
  fig.retina = 3, dpi = 300, fig.align = "center"
)
```

## Application Exercise

::: appex
📋 AE 15 - Inference for MLR

- Complete Exercises 0-1
:::


## Topics

::: nonincremental
-   Inference for multiple linear regression
-   Effect Size and Power
:::

## Computational setup

```{r}
#| echo: true


# load packages
library(tidyverse)
library(broom)
library(mosaic)
library(Stat2Data)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)
library(GGally)
library(countdown)
library(rms)
library(coursekata)

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))
```


## Data: `Tadpoles` {.smaller}

::: nonincremental
Biologists wondered whether tadpoles can adjust the relative length of their intestines if they are exposed to a fungus called Batrachochytrium dendrobatidis (Bd).
:::

```{r}
#| echo: false
data("Tadpoles")
Tadpoles |> glimpse()
```


Source: [Stat2Data](https://cran.r-project.org/web/packages/Stat2Data/refman/Stat2Data.html#Tadpoles).



# Conduct a hypothesis test for $\beta_j$

## [Review: Simple linear regression (SLR)]{.r-fit-text}
```{r}
p1 <- gf_jitter(GutLength ~ Treatment, data = Tadpoles, alpha = 0.5, height = 0, width = 0.1) |>
  gf_model(lm(GutLength ~ Treatment, data = Tadpoles), color = "red") |>
  gf_labs(x = "Treatment", y = "Gut Length (mm)")
p2 <- gf_point(GutLength ~ MouthpartDamage, data = Tadpoles, alpha = 0.5) |>
  gf_model(lm(GutLength ~ MouthpartDamage, data = Tadpoles), color = "red") |>
  gf_labs(x = "Mouthpart Damage", y = "Gut Length (mm)")
p1 + p2
```



## SLR model summary

```{r}
#| code-fold: true
treatment_model <- lm(GutLength ~ Treatment, data = Tadpoles)
damage_model <- lm(GutLength ~ MouthpartDamage, data = Tadpoles)
tidy(treatment_model) |> kable()
tidy(damage_model) |> kable()
```


## SLR hypothesis test {.smaller}

1.  **Set hypotheses:** $H_0: \beta_1 = 0$ vs. $H_A: \beta_1 \ne 0$

. . .

2.  **Calculate test statistic and p-value**
. . .

3.  **State the conclusion**

## Multiple linear regression

- Discussion: How should we handle our confounding variables?

. . .

Complete Exercise 2.

## Multiple linear regression

The multiple linear regression model assumes $$Y|X_1, X_2,  \ldots, X_p \sim N(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \dots + \beta_p X_p, \sigma_\epsilon^2)$$

. . .

For a given observation $(x_{i1}, x_{i2}, \ldots, x_{ip}, y_i)$, we can rewrite the previous statement as

$$y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \dots + \beta_p x_{ip} + \epsilon_{i}, \hspace{10mm} \epsilon_i \sim N(0,\sigma_{\epsilon}^2)$$


## Estimating $\sigma_\epsilon$ {.smaller}

For a given observation $(x_{i1}, x_{i2}, \ldots,x_{ip}, y_i)$ the residual is
$$
\begin{aligned}
e_i &= y_{i} - \hat{y_i}\\
&= y_i - (\hat{\beta}_0 + \hat{\beta}_1 x_{i1} + \hat{\beta}_{2} x_{i2} + \dots + \hat{\beta}_p x_{ip})
\end{aligned}
$$

. . .

The estimated value of the regression standard error , $\sigma_{\epsilon}$, is

$$\hat{\sigma}_\epsilon  = \sqrt{\frac{\sum_{i=1}^ne_i^2}{n-p-1}}$$

. . .

As with SLR, we use $\hat{\sigma}_{\epsilon}$ to calculate $SE_{\hat{\beta}_j}$, the standard error of each coefficient. See [Matrix Form of Linear Regression](https://github.com/STA210-Sp19/supplemental-notes/blob/master/regression-basics-matrix.pdf) for more detail.

## MLR Model

```{r}
lm(GutLength ~ Treatment + Body, data = Tadpoles) |>
  tidy()
```

## MLR hypothesis test: treatment {.smaller}

1.  **Set hypotheses:** $H_0: \beta_{trmt} = 0$ vs. $H_A: \beta_{trmt} \ne 0$, given other `Body` is in the model

. . .

2.  **Calculate test statistic and p-value:** The test statistic is $t = 1.48$. The p-value is calculated using a $t$ distribution with $$(n - p - 1) = 27-2-1 = 24$$ degrees of freedom. The p-value is $\approx 0.153$.
  
    + Note that $p$ counts all non-intercept $\beta$'s

. . .

3.  **State the conclusion:** The p-value provides no evidence in favor of the alternative, so we fail to reject $H_0$. The data does not provide evidence that a exposure to the *Bd* fungus is a useful predictor *in a model that already contains body length as a predictor of gut length* (could also say *when controlling for body length*).


## MLR hypothesis test: interaction terms {.smaller}

- Framework: Same as previous slide except use $\beta$ of interaction term
- $p$ (the number of predictors) should include the interaction term
- Conclusion: tells you whether interaction term is a useful predictor
- Warning: if $X_1$ and $X_2$ have an interaction term, don't try to interpret their individual p-values... if interaction term is significant, then both variables are important

::: appex
Complete Exercises 3.
:::

# Effect Size and Power

## Overview {.smaller}

-   **Effect Size:** 
    +   General Idea: how big is your parameter compared to the variability of your data
    +   Why is it important: if parameter has small effect size, it may not be a useful parameter to include. However, what effect size is useful depends on your problem
    +   In practice: many formal ways of determining effects size which we won't talk about

-   **Statistical Power:** the ability of a hypothesis test to detect a given **effect size**
    +   more data = more power
    +   tons of data = tons of power
    +   In practice: think about how what effect size you'd like to be able to detect and design your study to collect enough data to do so

## Effect Size

-  A few measures of effect size
  +  Cohen's $d$ (for when you have two groups: e.g. treament vs. control)
  +  $R^2$ (for when you have continuous predictors)
-  When evaluating effect size, consider practical significance not just statistical significance
-  Effect size DOESN'T depend on sample size (other than sampling variability)
-  Statistical significance DOES depend on sample size

## Power

-   Power: probability of correctly rejecting null hypothesis given a certain effect size
-   Power depends on:
    +   Effect size
    +   Sample size
    +   Significance level ($\alpha$)
    +   Variability in the data
-   In practice: decide on a minimum effect size you want to be able to detect, then design your study to have enough power to detect that effect size

## Power Analysis using Simulation

```{r}
lm(GutLength ~ Treatment + Body, data = Tadpoles) |>
  tidy()
```

## Power Analysis using Simulation I

```{r}
set.seed(2025)
sample_size <- 50
lm(GutLength ~ Treatment + Body, data = resample(Tadpoles, sample_size)) |>
  tidy()
```

## Power Analysis using Simulation II

```{r}
sample_size <- 50
lm(GutLength ~ Treatment + Body, data = resample(Tadpoles, sample_size)) |>
  tidy()
```

## Power Analysis using Simulation III

```{r}
sample_size <- 50
do(10) * lm(GutLength ~ Treatment + Body, data = resample(Tadpoles, sample_size)) |>
  tidy()
```

## Power Analysis using Simulation IV

```{r}
sample_size <- 50
do(10) * lm(GutLength ~ Treatment + Body, data = resample(Tadpoles, sample_size)) |>
  tidy() |>
  filter(term == "TreatmentControl")
```

## Power Analysis using Simulation IV

```{r}
sample_size <- 50
samp_dist <- do(1000) * lm(GutLength ~ Treatment + Body, data = resample(Tadpoles, sample_size)) |>
  tidy() |>
  filter(term == "TreatmentControl")

tally(samp_dist$p.value < 0.05, format = "proportion")
```

The power for a sample of size 50 if we assume the same effect size as in our data is approximately 0.56.

Complete Exercise 4.

# Confidence interval for $\beta_j$

## Confidence interval for $\beta_j$ {.midi}

-   The $C\%$ confidence interval for $\beta_j$ $$\hat{\beta}_j \pm t^* SE(\hat{\beta}_j)$$ where $t^*$ follows a $t$ distribution with $n - p - 1$ degrees of freedom.

-   **Generically**: We are $C\%$ confident that the interval LB to UB contains the population coefficient of $x_j$.

-   **In context:** We are $C\%$ confident that for every one unit increase in $x_j$, we expect $y$ to change by LB to UB units, *holding all else constant*.

::: appex
Complete Exercise 5.
:::

# Inference pitfalls

## Large sample sizes

::: callout-caution
If the sample size is large enough, the test will likely result in rejecting $H_0: \beta_j = 0$ even $x_j$ has a very small effect on $y$.

::: nonincremental
-   Consider the **practical significance** of the result not just the statistical significance.

-   Instead of saying **statistically significant** say **statistically detectable**

-   Use confidence intervals to draw conclusions instead of relying only on p-values.
:::
:::

## Small sample sizes

::: callout-caution
If the sample size is small, there may not be enough evidence to reject $H_0: \beta_j=0$.

::: nonincremental
-   When you fail to reject the null hypothesis, **DON'T** immediately conclude that the variable has no association with the response.

-   There may be a linear association that is just not strong enough to detect given your data, or there may be a non-linear association.
:::
:::


## Recap

- Inference for individual predictors
    + Hypothesis testing: same as SLR but add *if all other terms are in the model*
    + Confidence intervals: same as SLR but add *holding all other variables in the model constant*
- Effect size and power
    + Effect size: how big is your parameter compared to variability of data
    + Power: ability to detect a given effect size
>>>>>>> ed75e5b977868051a3347368172b222db58f174a
