---
title: "Multicollinearity"
author: "Prof. Eric Friedlander"
footer: "[🔗 MATH 2025 - Fall 2025 - Schedule](https://math2025fa25.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib  
---

```{r}
#| include: false

# figure options
knitr::opts_chunk$set(
  fig.width = 10, fig.asp = 0.618,
  fig.retina = 3, dpi = 300, fig.align = "center"
)
```



::: appex
📋 AE 17 - Multicollinearity

- Open up AE 17 and Complete Exercise 0
:::

## Topics

::: nonincremental
-   Defining Multicollinearity
-   Detecting Multicollinearity
-   Variance Inflation Factors
:::

## Computational setup

```{r}
#| echo: true


# load packages
library(tidyverse)
library(broom)
library(mosaic)
library(mosaicData)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)
library(countdown)
library(rms)

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))
```


## Data: `rail_trail` {.smaller}

::: nonincremental
-   The Pioneer Valley Planning Commission (PVPC) collected data for ninety days from April 5, 2005 to November 15, 2005.
-   Data collectors set up a laser sensor, with breaks in the laser beam recording when a rail-trail user passed the data collection station.
:::

```{r}
#| echo: false
rail_trail <- read_csv("../data/rail_trail.csv")
rail_trail
```


Source: [Pioneer Valley Planning Commission](http://www.fvgreenway.org/pdfs/Northampton-Bikepath-Volume-Counts%20_05_LTA.pdf) via the **mosaicData** package.

## Full model

```{r}
#| echo: FALSE
rt_full_fit <- lm(volume ~ ., data = rail_trail)
tidy(rt_full_fit) |> kable()

rt_full_aug <- augment(rt_full_fit)
```

# Multicollinearity

## What is multicollinearity

-   **Multicollinearity** is the case when one or more predictor variables are strongly correlated with some combination of other predictors
-   **Intuition:** if you could fit a good linear model with one of your predictors as the response and the rest of the predictors as your explanatory variables, then your predictors are exhibiting multicollinearity



## Example

Let's assume the true population regression equation is $y = 3 + 4x$

. . .

Suppose we try estimating that equation using a model with variables $x$ and $z = x/10$

$$
\begin{aligned}\hat{y}&= \hat{\beta}_0 + \hat{\beta}_1x  + \hat{\beta}_2z\\
&= \hat{\beta}_0 + \hat{\beta}_1x  + \hat{\beta}_2\frac{x}{10}\\
&= \hat{\beta}_0 + \bigg(\hat{\beta}_1 + \frac{\hat{\beta}_2}{10}\bigg)x
\end{aligned}
$$

## Example {.smaller}

$$\hat{y} = \hat{\beta}_0 + \bigg(\hat{\beta}_1 + \frac{\hat{\beta}_2}{10}\bigg)x$$

-   We can set $\hat{\beta}_1$ and $\hat{\beta}_2$ to any two numbers such that $\hat{\beta}_1 + \frac{\hat{\beta}_2}{10} = 4$

-   Therefore, we are unable to choose the "best" combination of $\hat{\beta}_1$ and $\hat{\beta}_2$

-   In statistics, we say this model is "unidentifiable" because different parameters combinations can result in the same model

-   This is also why we need to set a reference level for categorical variables

-   Complete Exercises 1-2.

## Why multicollinearity is a problem

-   When we have perfect collinearities, we are unable to get estimates for the coefficients

    -   When we have almost perfect collinearities (i.e. highly correlated predictor variables), the standard errors for our regression coefficients inflate

    -   In other words, we lose precision in our estimates of the regression coefficients

    -   This impedes our ability to use the model for inference

    -   It is also difficult to interpret the model coefficients

## Detecting Multicollinearity {.midi}

Multicollinearity may occur when...

-   There are very high correlations $(r > 0.9)$ among two or more predictor variables, especially when the sample size is small

-   One (or more) predictor variables can easily be predicted from our other predictor variables using linear regression

-   There are interactions between two or more continuous variables


## Detecting multicollinearity in the EDA

-   Look at a correlation matrix of the predictor variables, including all indicator variables
    -   Look out for values close to 1 or -1
-   Look at a scatter plot matrix of the predictor variables
    -   Look out for plots that show a relatively linear relationship
-   Complete Exercises 3-4.

## Detecting Multicollinearity (VIF)

**Variance Inflation Factor (VIF)**: Measure of multicollinearity in the regression model

$$VIF(\hat{\beta}_j) = \frac{1}{1-R^2_{X_j|X_{-j}}}$$

where $R^2_{X_j|X_{-j}}$ is the proportion of variation in $X_j$ that is explained by the other explanatory variables in the model.

## Detecting Multicollinearity (VIF)

-   Typically $VIF > 10$ indicates concerning multicollinearity

-   Variables with similar values of VIF are typically the ones correlated with each other

-   Use the `vif()` function in the **rms** R package to calculate VIF

## VIF for rail trail model

Complete Exercise 5.

. . .

```{r echo = T}
vif(rt_full_fit)
```

<br>

. . .

`hightemp` and `avgtemp` are correlated. 

## What to do about Multicollinearity

1. Drop some predictors.
    - Example: Remove <u>**one**</u> of these variables and refit the model.
2. Combine some predictors.
    - Example: Create a new variable `temp_comsite` that is the average of `avgtemp` and `hightemp`.
3. Discount the individual coefficients and t-tests.
    - Example: Think about `avgtemp` and `hightemp` together with their individual $\beta$'s and p-values not having much meaning.
    
. . .

Complete Exercises 6 & 7.

## Recap

-   Multicollinearity occurs when one or more predictor variables are strongly correlated with some combination of other predictors
-   Multicollinearity can inflate the standard errors of the regression coefficients, making inference difficult
-   Multicollinearity can be detected using correlation matrices, scatter plot matrices, and Variance Inflation Factors (VIF)
-   Possible solutions to multicollinearity include dropping predictors, combining predictors, or discounting individual coefficients and t-tests
