---
title: "Logistic regression continued"
subtitle: "Introduction"
author: "Prof. Eric Friedlander"
date: "2024-11-06"
date-format: "MMM DD, YYYY"
footer: "[🔗 MAT 212 - Fall 2024 -  Schedule](https://mat212fa24.netlify.app/schedule)"
logo: "../images/logo.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: false
    incremental: false 
    chalkboard: true
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
execute:
  freeze: auto
  echo: true
  cache: false
knitr:
  opts_chunk: 
    R.options:      
    width: 200
bibliography: references.bib
---

```{r}
#| include: false

# figure options
knitr::opts_chunk$set(
  fig.width = 10, fig.asp = 0.618, out.width = "90%",
  fig.retina = 3, dpi = 300, fig.align = "center"
)

library(countdown)
```

## Announcements

-   Project: [Paper Due November 18th](https://mat212fa24.netlify.app/project-instructions#draft-report)
-   Oral R Quiz

::: appex
📋 [AE 19 - Intro to Logistic Regression Continued](https://mat212fa24.netlify.app/ae/ae-19-logistic-intro2)

- Open up AE 19 and complete Exercise 0.
:::


# Logistic regression

## Topics


-   Use R to fit logistic regression models
-   Interpret coefficients of logistic regression models
-   Use logistic regression model to calculate predicted odds and probabilities

## Computational setup

```{r}
#| warning: false

# load packages
library(tidyverse)
library(ggformula)
library(ggforce)
library(broom)
library(knitr)

# set default theme and larger font size for ggplot2
ggplot2::theme_set(ggplot2::theme_bw(base_size = 20))
```


## Data: Framingham Study {.smaller}

This data set is from an ongoing cardiovascular study on residents of the town of Framingham, Massachusetts. We want to use the total cholesterol to predict if a randomly selected adult is high risk for heart disease in the next 10 years.

```{r}
heart_disease <- read_csv("data/framingham.csv") |>
  select(totChol, TenYearCHD, age, BMI, cigsPerDay, heartRate) |>
  drop_na()
```

## Variables

- Response: 
    -   `TenYearCHD`:
        -   1: Patient developed heart disease within 10 years of exam
        -   0: Patient did not develop heart disease within 10 years of exam
- Predictors: 
  -   `totChol`: total cholesterol (mg/dL)
  -   `age`: age in years


# Logistic regression

## Logistic regression model

**Logit form**: $$\log\big(\frac{\pi}{1-\pi}\big) = \beta_0 + \beta_1~X$$

. . .

**Probability form**:

$$
\pi = \frac{\exp\{\beta_0 + \beta_1~X\}}{1 + \exp\{\beta_0 + \beta_1~X\}}
$$

. . .

Today: Using R to fit this model.



## High risk vs. age

```{r}
#| echo: true

heart_disease |> 
gf_sina(age ~ factor(TenYearCHD)) |> 
  gf_labs(x = "High risk - 1: yes, 0: no",
       y = "Age", 
       title = "Age vs. High risk of heart disease")
```

## High risk vs. age

```{r}
#| echo: true

heart_disease |> 
gf_violin(age ~ factor(TenYearCHD), fill = "steelblue") |> 
  gf_labs(x = "High risk - 1: yes, 0: no",
       y = "Age", 
       title = "Age vs. High risk of heart disease")
```

## High risk vs. age

```{r}
#| echo: true

heart_disease |> 
gf_boxplot(age ~ factor(TenYearCHD), fill = "steelblue") |> 
  gf_sina(size = 0.75, alpha=0.25) |> 
  gf_labs(x = "High risk - 1: yes, 0: no",
       y = "Age", 
       title = "Age vs. High risk of heart disease")
```

## Let's fit a model

```{r}
#| echo: true
#| code-line-numbers: "1|3"
heart_disease_fit <- glm(TenYearCHD ~ age, data = heart_disease, family = "binomial")

tidy(heart_disease_fit) |> kable(digits = 3)
```

## The model

```{r}
#| echo: true

tidy(heart_disease_fit) |> kable(digits = 3)
```


$$\textbf{Logit form:}\qquad\log\Big(\frac{\hat{\pi}}{1-\hat{\pi}}\Big) = -5.561 + 0.075 \times \text{age}$$ 

$$\textbf{Probability form:}\qquad\hat{\pi} = \frac{\exp(-5.561 + 0.075 \times \text{age})}{1+\exp(-5.561 + 0.075 \times \text{age})}$$

where $\hat{\pi}$ is the predicted probability of developing heart disease in the next 10 years.

## Interpreting $\hat{\beta}$'s

```{r}
tidy(heart_disease_fit) |> kable(digits = 3)
```

For every addition year of age, the **log-odds** of developing heart disease in the next 10 years, increases by 0.077.

. . .


Complete Exercises 1 - 3.

## Interpretability of $\beta$ for predicted probabilities

```{r}
#| echo: false

augment(heart_disease_fit) |> 
  mutate(predicted_prob = exp(.fitted)/(1+exp(.fitted))) |> 
  gf_point(predicted_prob ~ age) |> 
  gf_labs( x = "Age", y = "Predicted Probability")
```

-   SLOPE IS CHANGING!
-   Increase in $\hat{\pi}$ due to increase of 1 year of Age *depends on what starting age is*

## `glm` and `augment` {.smaller}

The `.fitted` values in `augment` correspond to predictions from the logistic form of the model (i.e. the log-odds):

```{r}
augment(heart_disease_fit)  |> head()
```

. . .

Note: The residuals do not make sense here!

**For observation 1**

$$\text{predicted probability} = \hat{\pi} = \frac{\exp\{-2.650\}}{1 + \exp\{-2.650\}} = 0.066$$

## Using `predict` with `glm`

Default output is log-odds:

```{r}
predict(heart_disease_fit, new_data = heart_disease) |> head() |> kable(digits = 3)
```

## Using `predict` with `glm`

More commonly you want the predicted probability:

```{r}
predict(heart_disease_fit, newdata = heart_disease, type = "response") |> head() |> kable(digits = 3)
```

. . .

Complete Exercise 4


## Recap

-   Fit logistic regression models using `glm`
-   Interpreted coefficients in logistic regression models
-   Used logistic regression model to calculate predicted odds and probabilities
-   Use `predict` to make predictions using `glm`
