---
title: "HW 05: County Health"
subtitle: "Model Comparison/Evaluation + Outliers"
format:
  html:
    embed-resources: true
---

In this homework, you'll use simple and multiple linear regression to analyze the relationship between the number of doctors in a county, the number of beds, and the number of hospitals.

# Learning goals

By the end of the homework you will be able to...

-   compare models using $R^2$ and RMSE
-   transform observations to improve model fit
-   talk about outliers
-   fit a model with two predictors

# Getting started

- Go to [RStudio](https://rstudio.collegeofidaho.edu) and login with your College of Idaho Email and Password.

- Make a subfolder in your `hw` directory to store this homework.

- Log into [Canvas](https://cofi.instructure.com/courses/17093/assignments/202863), navigate to Homework 5 and upload the `hw-05.qmd` file into the folder your just made.

# Packages

We'll use the following packages in this homework.

```{r load-packages}
#| message: FALSE
#| eval: TRUE

library(tidyverse)
library(broom)
library(yardstick)
library(ggformula)
library(knitr)
library(patchwork)
library(Stat2Data)
library(GGally)
# add more packages as needed
```

## Data: County Health

The data set for this homework is from the `Stat2Data` R package which is the companion package for this course's textbook. It is the same data set that we used in AE-08. The data was originally generated by the American Medical Association and concerns the availability of health care in counties in the United States. You can find information [here](https://cran.r-project.org/web/packages/Stat2Data/Stat2Data.pdf) by searching for the County Health Resources dataset.

```{r}
#| label: load-data
#| warning: false
#| message: false

data("CountyHealth") # Loads the data from the package
```

It is relatively easy to count the number of hospitals a county has, whereas counting the number of doctors is much more difficult. We'd like to build a linear model to predict the number of doctors, contained in the variable `MDs`, from the number of hospitals, `Hospitals` and the number of beds, `Beds`.

## Exercise 1

Describe what an observational unit represents for this data set. How many are there?

## Exercise 2

In Example 1.7 of Stat2, they consider a simple linear model to predict the number of doctors (`MDs`) from the number of hospitals (`Hospitals`) in a metropolitan area. In that example, they found that a square root transformation on the response variable produced a more linear relationship. Create a new variable in the `CountyHealth` data frame called `sqrtMDs`. Hint: use the `sqrt` function inside the `mutate` function.

## Exercise 3

Use the function `ggpairs` from the package ``GGally`` to generate a grid scatter plots and correlations. Note that you will need to select the variables you want to use. Which explanatory variable (`Hospitals` or `Beds`) has the highest correlation with `SqrtMDs`?  Is this consistent with your visual assessment?


## Exercise 4

Fit a simple linear model using `SqrtMDs` as the response variable and `Hospitals` as the predictor. You may use `sqrt(MDs)` in your `lm` call instead of `SqrtMDs` if you like. How much of the variability in the `SqrtMDs` values is explained by `Hospitals`? How much of the variability in `MDs` is explained by the model you just fit. To figure this out:

1. Augment your model.
2. Convert the fitted and observed response variables back to number of MDs rather than square-root of the number of MDs.
3. Compute the $R^2$.

Why are these two numbers different?

## Exercise 5

Do you think taking the square root of `Beds` would improve this model? Support your argument with plots and/or numbers.

## Exercise 6

Repeat exercise 4 above with `Beds` as the predictor instead of `Hospitals`.

## Exercise 7

For the model you just fit, are there any high-influence outliers? Justify your answer using something from the lecture on outliers. There appear to be at least two high-leverage points. Which observations are these?

## Exercise 8

Fit a multiple linear model using `SqrtMDs` as the response variable and both `Hospitals` and `Beds` as the predictors. Interpret both slopes and the intercept in the context of the problem.

## Exercise 9

How much of the variation in `MDs` is explained by the model you just fit? Which model would you say is the "best", given what we've learned through the first lecture on multiple linear regression.

## Exercise 10

Using the "best" model, predict the average number of doctors in a metro area with 1,000 beds and 4 hospitals. Report a 95% prediction interval and interpret your results in context.


## Grading

Total points available: 20 points.

| Component             | Points |
|-----------------------|--------|
| Ex 1                  | 1      |
| Ex 2                  | 1      |
| Ex 3                  | 2      |
| Ex 4                  | 2      |
| Ex 5                  | 2      |
| Ex 6                  | 2      |
| Ex 7                  | 2      |
| Ex 8                  | 2      |
| Ex 9                  | 2      |
| Ex 10                 | 2      |
| Grammar & Writing     | 1[^1]  |
| Workflow & formatting | 1[^2]  |

[^1]: The "Grammar & Writing" grade is decided based on your grammar and writing. This is typically decided by choosing one of the questions and assessing the writing.

[^2]: The "Workflow & formatting" grade is to assess the reproducible workflow and document format. This includes having a neatly organized document with readable code and your name and the date in the YAML.
