---
title: "Homework 06: Candy Competition"
subtitle: "Inference for Multiple Linear Regression"
editor: visual
format:
  html:
    embed-resources: true
---

## Introduction

In today's homework you will analyze data about candy that was collected from an online experiment conducted at [FiveThirtyEight](https://fivethirtyeight.com/).

### Learning goals

By the end of the homework you will be able to

-   Fit a linear model with multiple predictors and an interaction term
-   Fit a linear model with categorical predictors
-   Conduct inference on multiple linear models

## Getting started

- Go to [RStudio](https://rstudio.collegeofidaho.edu) and login with your College of Idaho Email and Password.

- Make a subfolder in your `hw` directory to store this homework.

- Log into [Canvas](https://cofi.instructure.com/courses/17093/assignments/202864), navigate to Homework 6 and upload the `hw-06.qmd` file into the folder your just made.

## Packages

The following packages are used in the lab.

```{r}
#| label: load-packages
#| message: false
#| eval: true

library(tidyverse)
library(broom)
library(ggformula)
library(fivethirtyeight)
library(knitr)
library(yardstick)

# add other packages as needed
```

## Data: Candy

The data from this lab comes from the the article FiveThirtyEight [*The Ultimate Halloween Candy Power Ranking*](https://fivethirtyeight.com/features/the-ultimate-halloween-candy-power-ranking) by Walt Hickey. To collect data, Hickey and collaborators at FiveThirtyEight set up an experiment people could vote on a series of randomly generated candy matchups (e.g. Reese's vs. Skittles). Click [here](http://walthickey.com/2017/10/18/whats-the-best-halloween-candy/) to check out some of the match ups.

The data set contains the characteristics and win percentage from 85 candies in the experiment. The variables are

| Variable           | Description                                                               |
|--------------------|---------------------------------------------------------------------------|
| `chocolate`        | Does it contain chocolate?                                                |
| `fruity`           | Is it fruit flavored?                                                     |
| `caramel`          | Is there caramel in the candy?                                            |
| `peanutalmondy`    | Does it contain peanuts, peanut butter or almonds?                        |
| `nougat`           | Does it contain nougat?                                                   |
| `crispedricewafer` | Does it contain crisped rice, wafers, or a cookie component?              |
| `hard`             | Is it a hard candy?                                                       |
| `bar`              | Is it a candy bar?                                                        |
| `pluribus`         | Is it one of many candies in a bag or box?                                |
| `sugarpercent`     | The percentile of sugar it falls under within the data set. Values 0 - 1. |
| `pricepercent`     | The unit price percentile compared to the rest of the set. Values 0 - 1.  |
| `winpercent`       | The overall win percentage according to 269,000 matchups. Values 0 - 100. |

Use the code below to get a glimpse of the `candy_rankings` data frame in the **fivethirtyeight** R package.

```{r}
glimpse(candy_rankings)
```

## Exercises

The goal of this analysis is to use multiple linear regression to understand the factors that make a good candy.

## Exercise 1

Notice that the values of `pricepercent` and `sugarpercent` are proportions. User `mutate` to change the scale so that they are percentages.

## Exercise 2

-   Our response variable in this homework will be `winpercentage`. Choose two additional variables, one quantitative and one categorical. Generate a SINGLE plot that visualizes all three variables. Hint: remember that you can tie any aesthetic in your plot (e.g. color) to a variable be writing `aesthetic = ~variable_name`. For example `color = ~variable1` will color the points based on `variable1`.
-   Write two observations from your plot.

## Exercise 3

Fit a linear model including both variables you chose above.

## Exercise 4

Interpret the following in the context of the data:

-   Intercept

-   Coefficient of your quantitative variable

-   Coefficient(s) of your categorical variables

## Exercise 5

Choose one of the coefficients from your model and write out all the steps in the hypothesis test that corresponds to it's p-value as in [this slide](https://mat212wi25.netlify.app/slides/08-mlr-inference#/mlr-hypothesis-test-avgtemp).

## Exercise 6

Interpret the other p-values in your model in the context of the data. You do not need to write out the full hypothesis testing framework as you did in Exercise 5.

## Exercise 7

Generate 95% confidence intervals for all of the slopes in your model and interpret them in the context of the data.

## Exercise 8

Now fit a model using the same explanatory variables, but include an interaction term. Interpret the coefficient of your interaction term in the context of the data.

## Exercise 9

Interpret the p-value associated with the interaction term.


## Exercise 10

Based on the results from exercise 9, would you choose the model with or without the interaction term?

## Grading

Total points available: 15 points.

| Component             | Points |
|-----------------------|--------|
| Ex 1                  | 1      |
| Ex 2                  | 2      |
| Ex 3                  | 1      |
| Ex 4                  | 1      |
| Ex 5                  | 1      |
| Ex 6                  | 1      |
| Ex 7                  | 2      |
| Ex 8                  | 2      |
| Ex 9                  | 1      |
| Ex 10                 | 1      |
| Grammar & Writing     | 1[^1]  |
| Workflow & formatting | 1[^2]  |

[^1]: The "Grammar & Writing" grade is decided based on your grammar and writing. This is typically decided by choosing one of the questions and assessing the writing.

[^2]: The "Workflow & formatting" grade is to assess the reproducible workflow and document format. This includes having a neatly organized document with readable code and your name and the date in the YAML.
