---
title: "AE 03a: Simulation-Based Inference (HT Focus)"
subtitle: "Bikeshare"
editor: visual
format:
  html:
    embed-resources: true
---

::: callout-important
-   Open [RStudio](https://rstudio.collegeofidaho.edu/) and create a subfolder in your AE folder called "AE-03a"
-   Upload the `ae-03a-sbi-ht.qmd` and `dcbikeshare.csv` files into the folder you just created.
:::

```{r load-packages}
#| message: false
library(tidyverse)
library(ggformula)
library(broom)
library(openintro)
library(kableExtra)
```

## Data

Our dataset contains daily rentals from the Capital Bikeshare in Washington, DC in 2011 and 2012. It was obtained from the `dcbikeshare` data set in the `dsbox` R package.

We will focus on the following variables in the analysis:

-   `count`: total bike rentals
-   `temp_orig`: Temperature in degrees Celsius

```{r glimpse-data}
bikeshare <- read_csv("../data/dcbikeshare.csv") |>
  mutate(season = case_when(
    season == 1 ~ "winter",
    season == 2 ~ "spring",
    season == 3 ~ "summer",
    season == 4 ~ "fall"
  ),
  season = factor(season))
glimpse(bikeshare)
```

## Exercises

In this activity, each group will be assigned a season to use:

-   Groups 1 and 5: `winter`
-   Groups 2 and 6: `spring`
-   Groups 3 and 7: `summer`
-   Groups 4 and 8: `fall`

### Exercise 0

Filter the `bikeshare` data set so that it only contains observations from your assigned season. Make sure to give the new data set a different name.

### Exercise 1

Conduct a little EDA. Generate a scatter plot of the number of bike rentals vs the temperature for your season by filling in the blanks in the code below. What do you think `alpha` does?

```{r scatterplot}
#| eval: FALSE

gf______(______, data = ______, alpha = 0.7) |> 
  gf_labs(
    x = "Temperature (C)",
    y = "Number of Bike Rentals",
  )
```

### Exercise 2

Fit a simple linear regression model, display the results to two decimal places, and be prepared to discuss the interpretation of the resulting slope and intercept.

```{r fit-model}
#| eval: FALSE

model_fit <- lm(_____ ~ ______, data = _____)

tidy(_____)
```

### Exercise 3

Write down your research question in words then state the null and alternative hypotheses in both words and mathematical notation. You can use dollar signs to engage "math mode".

$$
H_0: 
$$

### Exercise 4

Generate null distribution using permutation

Fill in the code, then set `eval: true` .

```{r permutation}
#| eval: false

n = 100
set.seed(212)

null_dist <- _____ |>
  specify(______) |>
  hypothesize(null = "independence") |>
  generate(reps = _____, type = "permute") |>
  fit()
```

### Exercise 5

Visualize the null distribution. Does your slope seem unusual if we assume the null hypothesis is the truth?

```{r null-dist}
# Code for histogram of null distribution
```

### Exercise 6

Fill in the code below to compute the p-value. Have your reporter write the value on the board.

```{r calc-pvalue}
#| eval: false 

# get observed fit 
observed_fit <- _____ |>
  specify(____) |>
  fit()

# calculate p-value
get_p_value(
  ____,
  obs_stat = ____,
  direction = "two-sided"
)
```

### Exercise 7

-   Do larger or smaller p-values provide evidence for the alternative hypothesis?
-   Do larger or smaller p-values provide evidence for your research question?
-   Interpret your p-value in the context of the problem. Do you think your data provides strong evidence for your research question?

::: callout-important
To submit the AE:
- Render the document to produce the HTML with all of your work from today's class.
- The driver for your group should upload your .qmd and .html files to the Canvas assignment.
:::