---
title: "AE 05: Model Conditions"
subtitle: "Songs on Spotify"
editor: visual
format:
  html:
    embed-resources: true
---

::: callout-important
-   Open [RStudio](https://rstudio.collegeofidaho.edu/) and create a subfolder in your AE folder called "AE-05"

-   Go to the [Canvas](https://cofi.instructure.com/courses/17093/assignments/202834) and locate your `AE 05` assignment to get started.

-   Upload the `ae-05.qmd` and `spotify-popular.csv` files into the folder you just created.
:::

```{r}
#| label: load-packages
#| warning: false
#| message: false
library(tidyverse)
library(ggformula)
library(broom)
library(knitr)
library(patchwork) #arrange plots in a grid
```

## Introduction

This is a continuation of AE-04. The **Data** section below is the same as in that exercise. Feel free to skip it if you feel you remember everything about the data set or simply use it as a reference when needed.

## Data

The data set for this assignment is a subset from the [Spotify Songs](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md) Tidy Tuesday data set. The data were originally obtained from Spotify using the **spotifyr** R package.

It contains numerous characteristics for each song. You can see the full list of variables and definitions [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md). This analysis will focus specifically on the following variables:

| variable | class | description |
|:-----------------|:-----------------|:------------------------------------|
| track_id | character | Song unique ID |
| track_name | character | Song Name |
| track_artist | character | Song Artist |
| track_popularity | double | Song Popularity (0-100) where higher is better |
| danceability | double | Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable. |
| energy | double | Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy. |
| loudness | double | The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typical range between -60 and 0 db. |
| valence | double | A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry). |
| tempo | double | The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration. |
| duration_ms | double | Duration of song in milliseconds |

```{r}
#| label: load-data
#| warning: false
#| message: false
spotify <- read_csv("../data/spotify-popular.csv")
```

What makes a song danceable? To answer this question, we'll analyze data on some of the most popular songs on Spotify, i.e. those with `track_popularity >= 80`. We'll use linear regression to fit some models to predict a song's `dancebility`. Each group will be assigned a different predictor variable.

### Predictor Assignment:

Below is your assigned predictor. Look at the table above for a definition.

-   **Group 1:** `energy`
-   **Group 2:** `loudness`
-   **Group 3:** `valence`
-   **Group 4:** `tempo`

## Exercise 0

Fit a model using your assigned explanatory variable to predict a songs `danceability`. Use `tidy` and `kable` to neatly display your model and have your reporter write you model on the white board along with a confidence interval and p-value for the slope. Be prepared to discuss the interpretation of these quantities.

```{r fit-model}
## add code
```

## Exercise 1

Last time, we were too hasty in performing statistical inference. We need to make sure you data satifies the four conditions of inference. Let's check those model conditions now. Fill in the code below to use the `augment()` function to create a new data frame containing the residuals and fitted values (among other information)/

::: callout-important
Note: Remove `#|eval: false` from the code chunk after you have filled in the code.
:::

```{r}
#| eval: false
spotify_aug <- augment(_____)
```

## Exercise 3

Make a plot of the residual vs. fitted values.

```{r}
#| label: resid-vs-fitted

# add code here
```

## Exercise 4

Fill in the code to make a histogram of the residuals and a normal QQ-plot.

```{r}
#| label: resid-dist
#| eval: false

resid_hist <- gf_histogram(~____, data = ____)) |> 
  gf_labs(x = "_____", 
       y = "_____", 
       title = "____")

resid_qq <- gf_qq(~____, data = ____)  |>
  gf_qqline() |> 
  gf_labs(x = "_____", 
       y = "_____", 
       title = "____")

resid_hist + resid_qq
```

## Exercise 5

Assess the four model conditions. Use the plots from the previous exercises to help make the assessment.

-   Linearity
-   Constant variance
-   Normality
-   Independence

::: callout-important
To submit the AE:

-   Render the document to produce the HTML file with all of your work from today's class.
-   Upload your QMD and HTML files to the Canvas assignment.
:::
