---
title: "AE 17: Multicollinearity"
subtitle: "Rail Trail"
editor: visual
format:
  html:
    embed-resources: true
---

::: callout-important
-   Open [RStudio](https://rstudio.collegeofidaho.edu/) and create a subfolder in your AE folder called "AE-17".

-   Go to the [Canvas](https://cofi.instructure.com/courses/15692/assignments/184069) and locate your `AE-17` assignment to get started.

-   Upload the `ae-17.qmd` and `rail_trail.csv` files into the folder you just created. The `.qmd` and PDF responses are due in Canvas. You can check the due date on the Canvas assignment.
:::

## Packages + data

```{r load-packages}
#| message: false
library(tidyverse)
library(ggformula)
library(GGally)
library(broom)
library(knitr)
library(mosaic)
library(mosaicData)

rail_trail <- read_csv("../data/rail_trail.csv")
```

The data for this AE is based on data from the Pioneer Valley Planning Commission (PVPC) and is included in the **mosaicData** package. The PVPC collected data for ninety days from April 5, 2005 to November 15, 2005. Data collectors set up a laser sensor, with breaks in the laser beam recording when a rail-trail user passed the data collection station. More information can be found [here](http://www.fvgreenway.org/pdfs/Northampton-Bikepath-Volume-Counts%20_05_LTA.pdf).

## Analysis goal

The goals of this activity are to:

-   Detect and remedy multicollinearity in our data

## Exercise 0

Fit a linear model to predict `volume` from ALL of the other predictors. The resulting model is called the "full model".

## Exercise 1

Create a new data frame called `rail_trail_new` with two new columns, `hightemp_c` and `avgtemp_c`. These two new columns should store the `hightemp` and `avgtemp` measured in Celsius rather than Fahrenheit. Note that the formula to convert between Fahrenheit and Celsius is $C = (F-32)\times \frac{5}{9}$. Make sure `hightemp` and `avgtemp` are still in your data set.

## Exercise 2

Try to refit the full model on your data. Do you notice anything odd about your coefficients? If so, what do you think is happening?

## Exercise 3

Add a comment next to each line of code explaining what that line of code is doing.

```{r}
#| message: false

library(corrplot)

rail_trail |> # Take the rail_trail data set and then
  select(-volume) |> # What do I do
  select(where(is.numeric)) |> # What do I do
  cor() |>   # What do I do
  corrplot(method="number") # What do I do


rail_trail |> 
  select(-volume) |> # What do I do
  ggpairs()
```

## Exercise 4

Based on the output from the previous exercise, do you have reason to believe that our data is exhibiting multicollinearity?

## Exercise 5

Pipe the model you fit in Exercise 0 (i.e. not including the variable measured in Celcius) into the function `vif`. Are any of the vif's concerning? If so, which ones.

## Exercise 6

Fit two models:

-   `m1`: use all predictors except for `avgtemp`
-   `m2`: use all predictors except for `hightemp`
-   `m3`: average `avgtemp` and `hightemp` together and use it instead of either `avgtemp` and `hightemp`

## Exercise 7

Use `glance` to compute the $R^2_{adj}$, AIC, and BIC for both of these models. Which should we choose?

## To submit the AE

::: callout-important
-   Render the document to produce the PDF with all of your work from today's class.
-   Upload your QMD and PDF files to the Canvas assignment.
:::
