---
title: "Exam 2 Practice"
author: "MATH 2025"
format:
  pdf:
    pdf-engine: pdflatex
    documentclass: exam
    include-in-header:
      - text: |
          \newcommand{\E}{\mathbb{E}}
          \newcommand{\Var}{\text{Var}}
          \newcommand{\SD}{\text{SD}}
          \newcommand{\PP}{\text{P}}
          \newcommand{\bq}{\begin{questions}}
          \newcommand{\eq}{\end{questions}}
          \newcommand{\bp}{\begin{parts}}
          \newcommand{\ep}{\end{parts}}
          \newcommand{\bs}{\begin{subparts}}
          \newcommand{\es}{\end{subparts}}
---

**Answer the questions in the spaces provided. Show your work wherever practicable. You are permitted to use a calculator but not any other resources.**
 

The data set for this exam is a subset from the Spotify Songs Tidy Tuesday data set. The data were originally obtained from Spotify using the **spotifyr** R package. **It is the same data set we used for our activities.**

It contains numerous characteristics for each song. This analysis will focus specifically on the following variables:

| variable         | class     | description  |
|:------------|:------------|:----------------------------------------------|
| track_popularity | double    | Song Popularity (0-100) where higher is better |
| energy           | double    | Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy. |
| valence          | double    | A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).  |
| danceability |	double |	Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable. |
| playlist_genre	 | character |	The genre of the song (`edm`, `latin`, `pop`, `r&b`, `rap`, `rock`)

```{r}
#| label: load-data
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(ggformula)
library(mosaic)
library(broom)
library(knitr)
library(patchwork) #arrange plots in a grid

spotify <- read_csv("../data/spotify-popular.csv")
```


```{r}
spotify |> 
  select(track_popularity, energy, valence, playlist_genre) |> 
  head() |> 
  kable()
```

\newpage

For questions 1-7, consider `model1` below:

```{r}
model1 <- lm(danceability ~ valence + energy, data=spotify)
model1 |> tidy() |> kable(digits = 3)
```

\bq

\question[5] Use the values above to write down the equation that represents the regression model. Use the variable names to represent the variables rather than $Y$'s and $X$'s.

\vfill

\question[10] Give a one sentence interpretation (in context) of the first two numbers in the estimate column (i.e. 0.643 and 0.260).

\vfill

\newpage

\question[5] Which of the following represents the predicted `danceability` for a single observation that has a `valence` of 0.5 and an `energy` of 0.25?

\begin{choices}
\item $0.643 + 0.260 - 0.125 = 0.778$
\item $0.643 + 0.260\times 0.5 - 0.125\times 0.25 = 0.7418$
\item $29.694 + 9.693\times 0.25 - 3.410\times 0.5 = 30.4123$
\item $0.260\times 0.25 - 0.125\times 0.5 = 0.0025$
\item Not enough information
\end{choices}

\question[5] What is the residual of this estimated `danceability`?

\begin{choices}
\item $-0.240$
\item $0.143$
\item $0.375$  
\item Not enough information to say
\end{choices}

\question[5] Suppose you generate a 95% prediction interval for this prediction. If we change this prediction interval to a confidence interval (that is make a prediction for the average of your response rather than a single prediction) the width of the interval would:

\begin{choices}
\item increase
\item decrease
\item remain unchanged (approximately)
\end{choices}



\question[15] Suppose you use bootstrapping to generate a 95% confidence interval for the slope of `valence` in `model1`. For each of the following changes, indicate whether the width of your interval would increase, decrease, or approximately remain the same (keeping all else constant):

\bp

\item Increase sample size:

\begin{choices}
\item increase
\item decrease
\item remain unchanged (approximately)
\end{choices}

\item Increase confidence level:

\begin{choices}
\item increase
\item decrease
\item remain unchanged (approximately)
\end{choices}

\item Increase the number of bootstrapped replicates:

\begin{choices}
\item increase
\item decrease
\item remain unchanged (approximately)
\end{choices}

\ep

\newpage

\question[15] 
Consider a hypothesis test to determine whether there is a relationship between `valence` and `danceability` in `model1`:

\bp

\item State the null and alternative hypothesis in both words and symbols.\vfill
\item Give a rough definition of a p-value in the context of this problem.\vfill
\item What would be the result of the test in the context of the problem if $\alpha = 0.05$?\vfill

\ep

\newpage

For question 8, consider `model2` below:

```{r}
#| eval: FALSE
model2 <- lm(danceability ~ valence + energy + playlist_genre, data=spotify)
```

\question[5] Recall that `playlist_genre` has 6 distinct genre's. How many *additional* parameters (i.e. $\beta$'s) would `model2` include that `model1` would not?

\begin{choices}
\item 5
\item 6
\item 10
\item 12
\item 15
\item 18
\end{choices}

\newpage 
<!-- \question[5] Compared to `model1`, the $\hat{\beta}_i$'s corresponding to `valence` and `energy` would: -->

<!-- \begin{choices} -->
<!-- \item Increase -->
<!-- \item Decrease -->
<!-- \item Remain the same -->
<!-- \item Not enough information to say -->
<!-- \end{choices} -->

\question[10] For the constant variance assumption of a linear model, sketch a plot that would show that this assumption would be violated. Make sure to clearly label your axes so I know what the plot represents. Feel free to also describe what I should be looking for in words if you feel that your plot is not sufficiently clear.\vfill

\newpage 

The code below will be helpful for the rest of the exam.

```{r}
spotify <- spotify |> 
  mutate(pop_indicator = as.factor(ifelse(playlist_genre == "pop", "pop", "not pop")))
```


\question[8]

Consider the following model:

```{r}
model1 <- lm(danceability ~ energy*pop_indicator, data=spotify)

model1 |> tidy() |> kable()
```

This model contains two different lines. One for `pop` songs and one for `not pop`. Write down the equations of both lines. For partial credit you may just write down the equation of the model.

\bp

\item \texttt{pop} line:

\vfill

\item \texttt{not pop} line:

\vfill

\ep

\newpage

For questions 11-18, consider the following models:

```{r}
model1 <- lm(danceability ~ energy*pop_indicator, 
             data=spotify)
model2 <- lm(danceability ~ valence, data=spotify)
model3 <- lm(danceability ~ valence + energy + pop_indicator, data=spotify)
model4 <- lm(danceability ~ valence + track_popularity + energy*pop_indicator, data=spotify)
```

\question[5] Which of the following metrics is the LEAST useful for comparing these four models:

\begin{choices}
\item $R^2$
\item $R^2_{adj}$
\item AIC
\item BIC
\item $C_p$
\end{choices}

\question[5] For the answer you selected above, which of the following is true (select all that apply):

\begin{choices}
\item It is fine to use for comparing models with the same number of parameters.
\item Using it can lead to overfitting because increasing the model complexity will always lead to a ``better'' number.
\item We only use it when we are looking at a model with one predictor.
\item Even though it isn't useful for comparing models of different sizes, it is useful for describing a model's quality once a model is chosen.
\item It penalizes models for being more complex.
\end{choices}

\question[6] Which of the following models are nested in one another (select all that apply):

\begin{choices}
\item model1 in model2
\item model1 in model3
\item model1 in model4
\item model2 in model3
\item model2 in model4
\item model3 in model4
\end{choices}

\newpage

For questions 14 - 18 consider the following output in addition to the code at the beginning of the previous page:

```{r}
#| echo: FALSE

model_list <- list(model1, model2, model3, model4)
tibble(Model = c("model1", "model2","model3","model4")) |> 
  bind_cols(bind_rows(lapply(model_list, glance))) |> 
  select(-df, -df.residual, -nobs, -sigma, - statistic, -p.value, -deviance, -logLik) |> kable(digits = 3)
```

\question[2] Which model is "best" according to $R^2$?

\begin{choices}
\item model1
\item model2
\item model3
\item model4
\end{choices}

\question[2] Which model is "best" according to $R^2_{adj}$?

\begin{choices}
\item model1
\item model2
\item model3
\item model4
\end{choices}

\question[2] Which model is "best" according to AIC?

\begin{choices}
\item model1
\item model2
\item model3
\item model4
\end{choices}

\question[2] Which model is "best" according to BIC?

\begin{choices}
\item model1
\item model2
\item model3
\item model4
\end{choices}

\question[3] Which model do you think is "best"? If there are two candidate models select the one which is more "parsimonious"? Very briefly state your rationale.

\begin{choices}
\item model1
\item model2
\item model3
\item model4
\end{choices}


\eq

