---
title: "AE 11: Multiple Logistic Regression"
editor: visual
format:
  html:
    embed-resources: true
---

::: callout-important
-   Open [RStudio](https://rstudio.collegeofidaho.edu/) and create a subfolder in your AE folder called "AE-11".

-   Go to the [Canvas](https://cofi.instructure.com/courses/17093/assignments/202842) and locate your `AE-11` assignment to get started.

-   Upload the `ae-11.qmd` and `framingham.csv` files into the folder you just created.
:::

## Packages

```{r}
#| label: load-pkgs-data
#| message: false
 
library(tidyverse)
library(broom)
library(ggformula)
library(mosaic)
library(knitr)
```

### Data: Framingham study

This data set is from an ongoing cardiovascular study on residents of the town of Framingham, Massachusetts. We want to predict if a randomly selected adult is high risk for heart disease in the next 10 years.

### Response variable

-   `TenYearCHD`:
    -   1: Patient developed heart disease within 10 years of exam
    -   0: Patient did not develop heart disease within 10 years of exam

### What are my predictor variables?

Based on your group, use the following as your predictor variables.

-   Group 1:
    -   `BMI`: patient's body mass index
    -   `diabetes`: 1 if patient has diabetes, 0 otherwise
-   Group 2:
    -   `cigsPerDay`: number of cigarettes patient smokes per day
    -   `currentSmoker`: 1 if current smoker, 0 otherwise
-   Group 3:
    -   `sysBP`: systolic blood pressure (mmHg)
    -   `BPMeds`: 1 if they are on blood pressure medication, 0 otherwise
-   Group 4:
    -   `totChol`: total cholesterol (mg/dL)
    -   `prevalentHyp`: 1 if patient was hypertensive, 0 otherwise

## Exercise 0

Edit the code below to drop any missing values, convert the categorical predictors into factors, and select only the eight variables listed above along with `TenYearCHD`. Make sure to do those thing in an order that makes sense:

```{r}
heart_disease <- read_csv("../data/framingham.csv")
```

## Exercise 1

Fit a logistic regression model predicting `TenYearCHD` using your two assigned predictors. Write your resulting model on the board in any format you choose.

## Exercise 2

For both of your coefficients complete the madlibs below interpreting the value, you may need to edit a bit to make the final sentence grammatically correct:

Quantitative variable: When \[variable name\] increases by \[number\], the typical \[increase/decrease\] in \[quantity\] is \[number\], \[qualifying statement that makes this different than when we have one predictor\].

Quantitative variable: When \[variable name\] increases by \[number\], the typical odds-ratio is multiplied by a factor of \[number\], \[qualifying statement that makes this different than when we have one predictor\].

Categorical variable: For patients who are \[insert category\] at the time of examination, the typical \[increase/decrease\] in \[quantity\] is \[number\] vs patients \[insert opposite category\], \[qualifying statement that makes this different than when we have one predictor\].

Categorical variable: For patients who are \[insert category\] at the time of examination, the odds-ratio typically \[increases/decreases\] by a factor of \[number\] vs patients \[insert opposite category\], \[qualifying statement that makes this different than when we have one predictor\].

## Exercise 3

Compute p-values for both of your coefficients using `tidy`.

## Exercise 4

For both of your coefficients write the null and alternative hypotheses (in both symbols and words), the test statistics, and the p-values on the board (just pick them out from the `tidy` output). Then, complete the madlibs below interpreting the p-values, you may need to edit the final statements to make them grammatically correct:

Quantitative variable: The \[quantity\] is \[size\], so we \[do something to\] $H_0$. The data \[do/do not\] provide sufficient evidence that \[variable name\] is a statistically significant predictor of \[variable\], \[qualifying statement that makes this different than when we have one predictor\].

Categorical variable: The \[quantity\] is \[size\], so we \[do something to\] $H_0$. The data \[do/do not\] provide sufficient evidence that \[variable name\] is a statistically significant predictor of \[variable\], \[qualifying statement that makes this different than when we have one predictor\].

## Exercise 5

Use `tidy` to compute 99% confidence intervals for both of your coefficients. Write them on the board.

## Exercise 6

Interpret the the confidence intervals below:

**Log-odds interpretation (Quantitative):**

**Odds-ratio interpretation (Quantitative):**

**Log-odds interpretation (Categorical):**

**Odds-ratio interpretation (Categorical):**

## Submission

::: callout-important
To submit the AE:

-   Render the document to produce the HTML file with all of your work from today's class.
-   Upload your QMD and HTML files to the Canvas assignment.
:::
